### 2024-11-30 paper/paper.qmd
final DiD model with robustness checks

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Steven Li
thanks: "Code and data are available at: [https://github.com/stevenli-uoft/Toronto_BikeShare_Causality](https://github.com/stevenli-uoft/Toronto_BikeShare_Causality)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

\newpage
# Introduction
Context and motivation paragraph

Estimand paragraph

Results paragraph

Why it's important

Paper structure

# Data {#sec-data}
## Overview
Tools, data sources, and context

## Measurement
- Detailed explanation of how cycling infrastructure changes are measured and recorded
- Discussion of how ridership is measured through the Bike Share system
- Talk about how the data goes from people to paper/data

## Data Cleaning
Overview of data cleaning steps to reach final dataframe. Mention that a much more details data processing methodology is outlined in appendix.

**discuss using relative year for pre/treatment/post periods**

## Outcome Variables
- Monthly ridership counts
- Seasonally adjusted ridership

\vspace*{\fill}
\begin{center}
**Line graph of ridership over relative months, both raw and adjusted riderhsip**
\end{center}
\vspace*{\fill}

## Treatment Variables
- Treatment: Infrastructure changes (with timing)
- Include visualization of treatment and control groups.

\vspace*{\fill}
\begin{center}
**Bar graph of upgrades and installation over time**
\end{center}
\vspace*{\fill}

## Summary Statistics
\vspace*{\fill}
\begin{center}
**Table of treatment vs control group statistics, pre/post treatment**
\end{center}
\vspace*{\fill}

# Model
## Model justification
Why DiD model

## Assumptions
Assumptions on data and definitons

## Model set-up
Mathematical notation of model

## Robustness Checks
### Parallel Trends Test
- Methodology and justification

### Placebo Tests
- Methodology and justification

### Heterogeneity Analysis
- Define subgroups (Types of bikeways)
- Methodology and justification

# Results
## Parallel Trend Analysis
\vspace*{\fill}
\begin{center}
**Visualization of parallel trend test**
\end{center}
\vspace*{\fill}

- Statistics and discussion

## Difference-in-Difference Model Estimates
\vspace*{\fill}
\begin{center}
**Visualization of DiD model**
\end{center}
\vspace*{\fill}

- Primary results
- Discussion of statistical results

## Robustness Checks
### Placebo Test Results
- Present results and discuss implications

### Heterogenous Effects
- Compare effects across bikeway types
- Include figures or tables

# Discussion
## Model Results Interpretation

## Policy Implications and Recommendations

## Limitations and Future Research

\newpage

\appendix

# Appendix {.unnumbered}
# Idealized Survey Methodology
## Survey Objectives
Following Stantcheva's [-@Stantcheva2023] framework, we recognize that surveys are not merely for data colletion, but rather "creating the process that will generate the data." The goal of our idealized survey is to measure otherwise "invisible factors" regarding cycling infrastructures influence on bike share usage, specifically:

  > - Causal relationship between infrastructure changes and ridership
  > - Mental processes behind route selection with/without bike lanes
  > - Perception of safety and convenience
  > - Counterfactual behavior (would people use bike share if infrastructure improves)

Having this idealized survey would help us directly answer our research question by creating controlled variation in our sampling strategy, comparing treatment areas with new or upgraded bike lanes against control areas without changes. This design allows us to measure both first-stage effects of infrastructure changes on perceptions and intended behavior, and second-stage effects on actual ridership through follow-up surveys, while controlling for demographic and geographic factors. Through carefully designed questions about safety perceptions, route preferences, and time-value tradeoffs, we can identify the specific mechanisms through which infrastructure improvements affect ridership patterns. Additionally, our sampling of different user types enables analysis of heterogeneous effects across current, former, and potential users in various neighborhoods, providing a comprehensive understanding of how infrastructure changes influence Bike Share adoption.

## Implementation Strategy
### Sample Size and Target Audience
Our survey will be targeted towards: current and former (In-active user for over 6 months) Bike Share members, and non-bike share users living within 200 meters of Bike Share Station. Our target sample size will be 3,000 respondents:

- 1,000 treatment area residents
- 1,000 control area residents
- 1,000 current Bike Share users

### Recruitment Channels
- **Digital Targeting:** Instagram/Facebook/X (Twitter) ads geo-targeting users within 200 meters to a bike share station. This method will target our entire audience group, and allow us to differentiate between treatment and control groups.
- **Direct Outreach**: Email distribution or SMS notifications to current and former bike share users living within 200 meters of a Bike Share station, identified through Bike Share Toronto's database.
- **Incentive Structure**: \$20 Bike Share gift card for completed surveys, with a bonus \$10 for follow-up survey completion

### Budget Allocation
Our allocation of a \$100,000 budget, will be:

- $15,000 to digital targeting costs of placing ads on social media platforms.
- $10,000 to Bike Share Toronto to distribute Emails and SMS notifications, taking advantage of their existing digital platform and data.
- $75,000 to incentives, ensuring that all respondents receive their \$20 Bike Share Toronto gift card, and the bonus \$10 gift card.

## Survey Design Elements
**Opening Module**

"Welcome! Thank you for participating in our survey about bike lane infrastructure and its effects on Bike Share ridership in Toronto. This survey is being conducted as part of research to better understand how infrastructure upgrades influence biking habits.

Your responses will help shape future urban transportation decisions. The survey will take approximately 5 minutes to complete. As a thank-you for your time, you will receive a \$20 Bike Share Toronto gift card, with an additional \$10 gift card for completing a follow-up survey in 3 months.

Your participation is completely voluntary, and all responses will remain confidential and anonymous.

For any questions or concerns, contact us at: stevency.li@mail.utoronto.ca"

**Key Questions**

1. **Time-Based Preference**: "Consider two possible routes for a 15-minute trip. Route A: 15 minutes using regular streets with cars. Route B: 20-minutes using protected bike lanes. Which would you choose for a Bike Share trip?

  > - Route A (shorter, regular streets)
  > - Not sure
  > - Route B (longer, protected lanes)"

2. **Route Selection Experiment** (Map showing City of Toronto's neighbourhoods): "Which areas of Toronto do you typically commute through? Please type in the neighbourhood(s) in the text box below."

3. **Visual Choice Experiment** (Show side-by-side images of the same bike lane before and after upgrades) : "If you needed to travel along this route, which version would make you more likely to use Bike Share?

  > - Much more likely with the upgraded lane
  > - Somewhat more likely with the upgraded lane
  > - No difference
  > - Less likely with the upgraded lane"

4. **Behavioral Scenarios**: "Imagine a Bike Share station gets installed near your workplace, and the city adds protected bike lanes on your route to work. What would you do?

  > - Definitely try biking to work
  > - Might try biking occasionally
  > - Probably wouldn't change my commuting method
  > - Definitely wouldn't change my commuting method"

**Confirmation Message**

"Thank you for your time and participation..."Thank you for completing this survey! Your insights will help us better understand how bike infrastructure influences transportation choices in Toronto.

Your \$20 Bike Share Toronto gift card will be sent to your provided contact information within 5 business days. In approximately 3 months, we will send you an invitation for a brief follow-up survey, where you can earn an additional \$10 gift card.

For any questions about the survey or your gift card, please contact: stevency.li@mail.utoronto.ca"

**Idealized Survey — Opening Module can be found at**: [First Survey Link](https://docs.google.com/forms/d/e/1FAIpQLSei00OAr1SRCIorOniT8Wf9GTmZLOdtRSjEFobWrUVxDddRYw/viewform)

**Follow-up Module**

As recommended by Stantcheva [-@Stantcheva2023] for testing persistence of effects, we will implement a comprehensive three-month follow-up survey. The key objective is to assess whether the stated preferences and intentions from the initial survey translate into actual behavioral changes, while also measuring any infrastructure-induced effects on ridership patterns. To maximize response rates, we will offer an additional $10 gift card incentive and keep the follow-up survey brief (under 3 minutes).

The follow-up survey will focus on concrete behavioral measures rather than attitudes, asking respondents about their actual Bike Share usage patterns over the past three months, including:

- Frequency of Bike Share trips
- Use of routes with new or upgraded infrastructure
- Reasons for any deviations from stated intentions

To minimize attrition bias, which Stantcheva [-@Stantcheva2023] identifies as a key concern in follow-up surveys, we will:

- Send reminder notifications through multiple channels (email, SMS, app notifications)
- Track and analyze patterns of non-response
- Apply appropriate statistical corrections for differential attrition

This approach allows us to not only validate initial survey responses but also capture how seasonal changes and growing familiarity with new infrastructure might influence ridership patterns. The combination of initial intentions and realized behaviors provides a more robust foundation for establishing the causal relationship between infrastructure improvements and Bike Share adoption.

**Idealized Survey — Follow-up Module can be found at**: [Follow-up Survey Link](https://docs.google.com/forms/d/e/1FAIpQLSeFFbTQ2TPpQ96Nw2xq_aPBEO_AE3N3X6o5bG2JSYYVLZyYXg/viewform)

## Bias Mitigation Strategies
We implement several strategies to mitigate potential biases in our survey design. To reduce social desirability bias, we emphasize complete anonymity of responses, employ indirect questioning techniques, and carefully avoid leading questions about environmental benefits that might prompt respondents to present themselves as more environmentally conscious. Survey fatigue is addressed by keeping the total length to 5 minutes, and varying question formats to maintain engagement. Selection bias poses a particular challenge in transportation surveys, so we collect detailed non-response data, and carefully document and analyze attrition patterns. This approach to bias mitigation helps ensure our results accurately reflect the relationship between cycling infrastructure and Bike Share usage, rather than capturing respondent's desires to appear environmentally conscious or systematic differences between those who complete versus abandon the survey.

## Tradeoffs and Limitations
Our survey methodology faces several important limitations that could affect the validity of our results. First, self-selection and non-response bias may skew our sample toward respondents who are already interested in cycling or have strong opinions about infrastructure, while our incentive structure might attract particular demographic groups, potentially missing key segments of the population. Our reliance on hypothetical scenarios and stated preferences introduces additional uncertainty, as actual behavior may differ significantly from reported intentions, especially regarding route selection and infrastructure usage. Despite careful design to minimize social desirability bias, respondents may still feel pressure to express support for sustainable transportation initiatives or understate safety concerns. Furthermore, temporal limitations of our three-month follow-up period may inadequately capture seasonal variations in cycling behavior and long-term adaptation to infrastructure changes, while the results may not generalize well to other urban contexts or represent the full diversity of potential users. These limitations suggest that while our survey can provide valuable insights into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.

# Additional Tables & Figures

# Data Processing
## Spatial Filtering
## Treatment and Control Group Construction

# Model Details and Diagnostics

\newpage

# References

### 2024-12-02 paper/paper.qmd
final edits to paper, cleaned up scripts structure

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Beyond Infrastructure: Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
+subtitle: "A Difference-in-Differences Analysis of Toronto's Cycling Network, 2017-2023"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+abstract: "We examine the causal effect of cycling infrastructure improvements on Bike Share Toronto ridership by analyzing how different types of bikeways influence system usage between 2017-2023.
+Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that infrastructure improvements led to significant increases in ridership, with post-implementation effects reaching 133.1 additional monthly rides per bikeway, translating to approximately 300,000 additional annual rides across all treated segments.
+Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides - more than double the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides), while upgrades to existing infrastructure showed more immediate benefits than new installations.
+Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027.
+These results provide evidence-based guidance for cities working to encourage sustainable transportation through cycling infrastructure development."
+format:
+  pdf:
+    toc: true
+include-before: |
+  \newpage
 number-sections: true
@@
 library(tidyverse)
+library(dplyr)
+library(knitr)
+library(lubridate)
+library(ggplot2)
+library(arrow)
+library(kableExtra)
+library(fixest)
+library(modelsummary)
+
+analysis_data <- read_parquet(here::here("data/02-analysis_data/final_df.parquet"))
 ```
@@
 # Introduction
-Context and motivation paragraph
+Cycling infrastructure plays a crucial role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
+While protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
+This knowledge gap is particularly relevant for Toronto, where rapid expansion of cycling infrastructure has coincided with dramatic growth in the bike share network from 2,854 to over 6,850 stations between 2016 and 2020 [@toronto_bike_share_expansion].
+Understanding this relationship is critical as cities increasingly rely on bike share systems to complement public transit and provide sustainable transportation options.
 
-Estimand paragraph
+We estimate the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, examining how protected lanes, on-road lanes, and shared roadways influence system usage.
+Using a difference-in-differences design, we analyze both new installations and infrastructure upgrades while controlling for broader trends in cycling adoption.
 
-Results paragraph
+Our analysis reveals substantial increases in bike share ridership following infrastructure improvements.
+Post-implementation effects reach 133.1 additional monthly rides per bikeway (p < 0.01), translating to approximately 300,000 additional annual rides across all 186 treated segments - particularly notable as these improvements represent only 16% of the 1,191 bikeways in our study.
+Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides (p < 0.001) - more than double the effect of on-road lanes and triple that of shared roadways.
+These gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment.
+This suggests infrastructure improvements can substantially increase cycling activity even on popular routes.
 
-Why it's important
+These findings carry important implications for urban transportation policy.
+Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027 [@cyclingNetworkPlan] presents an opportunity to enhance cycling adoption through strategic implementation.
+Our results challenge conventional assumptions about prioritizing underserved areas, as improvements to high-traffic routes generated significant additional ridership, highlighting the complex trade-offs in infrastructure resource allocation.
 
-Paper structure
+The remainder of this paper proceeds as follows.
+@sec-data describes our data sources and measurement approach.
+@sec-model outlines our difference-in-differences methodology and identification strategy.
+@sec-results presents our main results and robustness checks.
+@sec-discussion discusses the implications of our findings and potential directions for future research.
+@sec-ideal-survey introduces an idealized survey methodology designed to directly address our research question.
+The appendices provide supplementary materials, including detailed data processing steps in @sec-data-processing and additional figures and analysis in @sec-additional-figures.
 
@@
 ## Overview
-Tools, data sources, and context
+This study draws on three datasets from the City of Toronto's Open Data Portal [@torontoOpenData]: the Cycling Network (Bikeways) data [@bikeWayData], and both Bike Share Ridership and Stations data [@bikeShareData].
+Together, these datasets provide a comprehensive view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023.
+All data cleaning and analysis was conducted using R [@RCoreTeam2023], with key packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], and `sf` [@sf] for spatial analysis.
 
-## Measurement
-- Detailed explanation of how cycling infrastructure changes are measured and recorded
-- Discussion of how ridership is measured through the Bike Share system
-- Talk about how the data goes from people to paper/data
+## Measurement and Limitations
+### Cycling Infrastructure Measurement
+The Cycling Network data captures two distinct types of infrastructure changes recorded by the City's Transportation Services division [@bikeWayData]:
+
+1. **New Installation**: The construction of new bikeway segments that expand the network's reach
+2. **Upgrades**: Modifications to existing segments that enhance their safety or functionality
+
+When either change occurs, the location, infrastructure type, and year of modification are documented.
+A key limitation of this measurement process is its temporal granularity - while the year of each change is recorded, the specific month and day are not, constraining our ability to pinpoint exact implementation dates.
+
+### Ridership and Stations Measurement
+Bike Share ridership is measured through an automated system that anonymously records every trip's start and end points, time, and associated station IDs. The measurement process follows these steps:
+
+1. A user unlocks a bike using their membership card, mobile app, or payment kiosks located at each station
+2. The system logs the start time and station information
+3. Upon return, the end time and station information are recorded
+4. This data is automatically transmitted to Bike Share Toronto's database
+
+The Public Bicycle System Company (PBSC) and City of Toronto also maintains detailed Bike Share stations records, documenting each station's location and capacity metrics [@bikeWayData].
+
+### Limitations and Potential Biases
+Our analysis faces two primary measurement challenges that could affect causal inference.
+First, the Bike Share system cannot track actual route choices between endpoints.
+To address this, we focus on trips originating from stations within 100 meters of bikeways, assuming proximity indicates usage.
+While this provides a reasonable proxy, it may introduce selection bias by excluding riders who use bikeways despite starting from more distant stations.
+
+Second, our measure of infrastructure changes may suffer from reverse causality bias.
+The city might prioritize upgrades for routes that already show high cycling traffic, potentially overstating the causal effect of infrastructure improvements on usage.
+We attempt to mitigate this through our difference-in-differences design and by examining pre-treatment trends, but this limitation should be considered when interpreting our results.
 
 ## Data Cleaning
-Overview of data cleaning steps to reach final dataframe. Mention that a much more details data processing methodology is outlined in appendix.
+The data cleaning process involved several key steps to create our final analysis panel dataset (detailed methodology available in Appendix @sec-data-processing):
 
-**discuss using relative year for pre/treatment/post periods**
+1. Spatial filtering to identify stations proximate to bikeways (within 100 meters)
+2. Extract relevant Bike Share trips based on filtered station locations
+3. Aggregating total monthly rides for each bikeway, adjusting for seasonality
+4. Creating treatment and control groups:
+
+   - **Treatment group**: Bikeways upgraded or constructed between 2019-2021
+   - **Control group **: Bikeways unmodified since before 2017
+
+A crucial methodological choice was the use of relative timing rather than calendar years.
+Due to bikeways having different upgrade or installation years, we align their treatment event year using a relative period.
+For treated bikeways, this creates a 60-month window: 24 months pre-treatment, 12 months during the treatment year, and 24 months post-treatment.
+Control bikeways receive randomly assigned pseudo-treatment years (2019, 2020, or 2021) to enable parallel analysis.
+This approach allows us facilitate direct comparison between treatment and control groups while accounting for the temporal distribution of infrastructure changes.
+
+@tbl-sample-data in Appendix @sec-additional-figures provides a sample of our processed dataset.
 
 ## Outcome Variables
-- Monthly ridership counts
-- Seasonally adjusted ridership
+Our primary outcome variable is seasonally-adjusted monthly ridership for each bikeway (`monthly_rides_adj`).
+@fig-avg-monthly-ridership reveals several important patterns in ridership behavior.
+The raw ridership (black line) displays pronounced seasonal fluctuations, with consistent peaks during summer months and troughs in winter, reflecting Toronto's strong cycling seasonality.
+After applying seasonal adjustments (blue line), these cyclical patterns are smoothed while preserving the underlying growth trends.
+Both treatment and control groups show a general upward trend in ridership over the observation period, suggesting broader growth in cycling activity across Toronto's network.
 
-\vspace*{\fill}
-\begin{center}
-**Line graph of ridership over relative months, both raw and adjusted riderhsip**
-\end{center}
-\vspace*{\fill}
+The notable unadjusted ridership activity during the treatment period (relative month 0) reflects both the granularity of our data and real-world implementation patterns.
+Since we observe only the year of infrastructure changes due to data limitations, all twelve months of the treatment year are assigned to relative month 0.
+This ridership activity captures both partial-year effects (as changes may occur at different points throughout the year) and continued system usage during implementation.
+
+```{r}
+#| label: fig-avg-monthly-ridership
+#| fig-cap: "Average ridership by relative months to treatment"
+#| echo: false
+#| warning: false
+#| message: false
+
+ridership_by_month <- analysis_data %>%
+  group_by(relative_month, treatment) %>%
+  summarise(
+    avg_rides = mean(monthly_rides, na.rm = TRUE),
+    avg_rides_adj = mean(monthly_rides_adj, na.rm = TRUE)
+  ) %>%
+  pivot_longer(
+    cols = c(avg_rides, avg_rides_adj),
+    names_to = "metric",
+    values_to = "rides"
+  )
+
+ggplot(ridership_by_month, aes(x = relative_month, y = rides, color = metric)) +
+  geom_line() +
+  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
+  labs(
+    x = "Months Relative to Treatment",
+    y = "Average Monthly Rides",
+    color = "Metric"
+  ) +
+  scale_color_manual(
+    values = c("grey10", "blue"),
+    labels = c("Raw Rides", "Seasonally Adjusted")
+  ) +
+  theme_minimal() +
+  theme(legend.position = "top",
+        panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
+        ) +
+  facet_wrap(~treatment, scales = "free_y", 
+             labeller = labeller(treatment = c("TRUE" = "Treatment", "FALSE" = "Control")))
+
+```
 
 ## Treatment Variables
-- Treatment: Infrastructure changes (with timing)
-- Include visualization of treatment and control groups.
+Our main treatment variable (`treatment`) indicates whether a bikeway received an infrastructure change (either an upgrade or new installation) during our study period 2019-2021.
+@fig-treatment-control-count presents the distribution of bikeways between treatment and control groups, showing a larger pool of control bikeways to enable robust comparison with treated segments.
 
-\vspace*{\fill}
-\begin{center}
-**Bar graph of upgrades and installation over time**
-\end{center}
-\vspace*{\fill}
+Appendix @sec-additional-figures provide a further breakdown on types of bikeways installed and upgraded throughout the years.
+@fig-bikeway_type_constructed_graph in the appendix shows consistent bikeway installations every year after 2010, with protected lanes most commonly installed.
+@fig-bikeway_type_upgraded_graph presents yearly bikeway upgrades, showing an increase during our treatment period (2019-2021), especially with protected bike lanes.
 
-## Summary Statistics
-\vspace*{\fill}
-\begin{center}
-**Table of treatment vs control group statistics, pre/post treatment**
-\end{center}
-\vspace*{\fill}
+```{r}
+#| label: fig-treatment-control-count
+#| fig-cap: "Total Bikeway Count in Control Vs. Treatment Group"
+#| echo: false
+#| warning: false
+#| message: false
 
-# Model
+installation_counts <- analysis_data %>%
+  select(bikeway_id, treatment, construct_year) %>%
+  distinct() %>%
+  group_by(treatment) %>%
+  summarise(count = n()) %>%
+  mutate(
+    group_label = ifelse(treatment, "Treatment", "Control")
+  )
+
+ggplot(installation_counts, 
+       aes(x = group_label, y = count, fill = group_label)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  scale_fill_manual(
+    values = c("Control" = "#1f77b4", "Treatment" = "#ff7f0e")
+  ) +
+  labs(
+    x = "Group",
+    y = "Number of Bikeways",
+    fill = "Group"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "top",
+        panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
+        )
+```
+
+To explore treatment effect heterogeneity, we examine two additional dimensions of infrastructure changes:
+
+- **Bikeway Type**:
+
+  > - *Protected Lanes*: Physically separated cycling infrastructure
+  > - *On-Road Lanes*: Dedicated but unprotected bike lanes
+  > - *Shared Roadways*: Designated shared-use routes
+
+- **Installation Type**:
+
+  > - *Upgrades*: Improvements to existing infrastructure
+  > - *New Installations*: Addition of new cycling routes
+
+@fig-infra-type reveals distinct infrastructure patterns between groups.
+While protected lanes dominate both treatment and control groups, comprising approximately 50% of bikeways in each, the distribution of other types differs notably.
+The treatment group shows a higher proportion of on-road lanes relative to shared roadways, suggesting a deliberate strategy of upgrading less protected infrastructure.
+This aligns with Toronto's broader policy goal of enhancing cyclist safety through infrastructure improvements [@toronto_council_agenda].
+
+@fig-treatment-type-total demonstrates a nearly even split in treatment approaches, with 98 new installations and 88 upgrades among the 186 treated bikeways.
+This balanced distribution enables us to examine whether infrastructure expansion (new installations) versus enhancement (upgrades) generates different effects on ridership patterns.
+Further categorization into bikeway types, and sub-treatment groups allow us to examine whether certain types of infrastructures and types of changes yield different impacts on ridership patterns.
+
+```{r}
+#| label: fig-infra-type
+#| fig-cap: "Infrastrucutre Type Distribution between Control and Treatment Group"
+#| echo: false
+#| warning: false
+#| message: false
+
+infra_counts <- analysis_data %>%
+  select(bikeway_id, treatment, bikeway_type) %>%
+  distinct() %>%
+  group_by(treatment, bikeway_type) %>%
+  summarise(count = n())
+
+ggplot(infra_counts, aes(x = bikeway_type, y = count, fill = treatment)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(
+    x = "Infrastructure Type",
+    y = "Number of Bikeways",
+    fill = "Group"
+  ) +
+  scale_fill_manual(
+    values = c("#1f77b4", "#ff7f0e"),
+    labels = c("Control", "Treatment")
+  ) +
+  theme_minimal() +
+  theme(legend.position = "top",
+        panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
+        )
+
+```
+
+```{r}
+#| label: fig-treatment-type-total
+#| fig-cap: "Number of New-Installations Vs Upgrades in Treatment Group"
+#| echo: false
+#| warning: false
+#| message: false
+
+treatment_type_counts <- analysis_data %>%
+  filter(treatment == TRUE) %>%
+  select(bikeway_id, sub_treatment_type) %>%
+  distinct() %>%
+  group_by(sub_treatment_type) %>%
+  summarise(count = n()) %>%
+  filter(sub_treatment_type != "No Treatment")
+
+ggplot(treatment_type_counts, aes(x = sub_treatment_type, y = count)) +
+  geom_bar(stat = "identity", fill = "#ff7f0e") +
+  labs(
+    x = "Sub-treatment Type",
+    y = "Number of Bikeways"
+  ) +
+  theme_minimal() +
+  theme(panel.grid.major = element_line(color = "gray85", linewidth = 0.35))
+
+```
+
+# Model {#sec-model}
+## Assumptions
+Our model implementation and interpretation rests on two key assumptions.
+First, we assume Bike Share rides that start within 100 meters of a bikeway indicate usage of that infrastructure, as cyclists typically choose the most direct route to their destination and are unlikely to travel significant distances to access a bikeway.
+Second, we treat the year of bikeway construction or upgrade as generating a sharp treatment effect, though we account for potential construction overlap by excluding six months of pre-treatment data in our parallel trend test to mitigate anticipatory affects.
+Our 24-month post-treatment window allows sufficient time to observe behavioral adaptation while minimizing contamination from other infrastructure changes.
+Additionally, our use of seasonally adjusted ridership helps isolate infrastructure effects from weather-related variation in cycling patterns, as Toronto's distinct seasonal climate significantly influences cycling behavior.
+
 ## Model justification
-Why DiD model
+To estimate the causal effect of bikeway infrastructure changes on Bike Share ridership, we employ a difference-in-differences (DiD) approach.
+This methodology is particularly well-suited for our research question as it allows us to exploit the temporal variation in bikeway upgrades and installations while controlling for unobserved confounders.
+The staggered implementation of bikeway improvements across Toronto from 2019 to 2021 provides a natural experiment setting where we can compare ridership changes between treated and untreated bikeways before and after infrastructure improvements.
 
-## Assumptions
-Assumptions on data and definitons
+Our analysis is ran in R [@RCoreTeam2023] using the `fixest` package [@fixest], which provides estimation methods for fixed effects models.
+This package is particularly suitable for our analysis as it handles multiple fixed effects efficiently and provides robust standard errors clustered at the bikeway level.
 
-## Model set-up
-Mathematical notation of model
+## Model Setup
+Following the framework of @cuborican2024econometrics, let $y_{it}$ represent the seasonally-adjusted monthly Bike Share rides for bikeway $i$ in month $t$. Our main DiD specification is:
+
+$$y_{it} = \alpha_i + \gamma_t + \sum_{p \in \{treatment,post\}} \beta_p (Treatment_i \times Period_{p,it}) + \epsilon_{it}$$
+
+where:
+
+- $\alpha_i$ represents bikeway fixed effects
+- $\gamma_t$ represents relative-month fixed effects
+- $Treatment_i$ is an indicator equal to 1 if bikeway $i$ received an upgrade or was newly installed during 2019-2021
+- $Period_{p,it}$ are indicators for treatment period and post-treatment period (The pre-treatment period serves as the reference category)
+- $\epsilon_{it}$ is the error term
+
+The coefficients of interest are $\beta_{treatment}$ and $\beta_{post}$, which capture the average treatment effects during and after the infrastructure changes, respectively.
 
 ## Robustness Checks
+
 ### Parallel Trends Test
-- Methodology and justification
+A crucial assumption of the DiD design is that treated and control bikeways would have followed parallel trends in the absence of treatment [@cuborican2024econometrics].
+While this counterfactual cannot be directly tested, we can examine pre-treatment trends to assess the plausibility of this assumption.
+We implement this test by estimating:
+
+$$y_{it} = \alpha_i + \gamma_t + \sum_{k=-24}^{-6} \delta_k (Treatment_i \times 1[t=k]) + \epsilon_{it}$$
+
+where $k$ indexes months relative to treatment.
+Due to data limitations that only specify the year of infrastructure changes, we conservatively exclude the six months immediately preceding treatment to account for potential construction overlap from the previous year that could affect ridership patterns.
+This approach helps isolate the true pre-treatment period from any anticipation effects or early construction impacts.
+The coefficients $\delta_k$ should not be statistically different from zero if the parallel trends assumption holds.
 
 ### Placebo Tests
-- Methodology and justification
+To further validate our results, we conduct a placebo test by randomly reassigning treatment status among bikeways while maintaining the original proportion of treated units.
+This approach helps ensure our findings are not driven by spurious correlations or systematic differences unrelated to the actual infrastructure changes.
+The specification mirrors our main DiD model but uses the randomly assigned treatment status:
+
+$$y_{it} = \alpha_i + \gamma_t + \sum_{p \in \{treatment,post\}} \beta_p (PlaceboTreatment_i \times Period_{p,it}) + \epsilon_{it}$$
 
 ### Heterogeneity Analysis
-- Define subgroups (Types of bikeways)
-- Methodology and justification
+We extend our analysis to examine treatment effect heterogeneity across different types of infrastructure changes.
+This investigation is crucial for understanding which interventions are most effective at increasing ridership.
+We estimate two separate models:
 
-# Results
-## Parallel Trend Analysis
-\vspace*{\fill}
-\begin{center}
-**Visualization of parallel trend test**
-\end{center}
-\vspace*{\fill}
+1. **Bikeway Type Effects**:
 
-- Statistics and discussion
+$$y_{it} = \alpha_i + \gamma_t + \sum_{p \in \{treatment,post\}} \sum_{j \in J} \beta_{pj} (Treatment_{ij} \times Period_{p,it}) + \epsilon_{it}$$
 
-## Difference-in-Difference Model Estimates
-\vspace*{\fill}
-\begin{center}
-**Visualization of DiD model**
-\end{center}
-\vspace*{\fill}
+where $J = \{$Protected Lanes, On-Road Lanes, Shared Roadways$\}$
 
-- Primary results
-- Discussion of statistical results
+2. **Installation Type Effects**:
+
+$$y_{it} = \alpha_i + \gamma_t + \sum_{p \in \{treatment,post\}} \sum_{k \in K} \beta_{pk} (Treatment_{ik} \times Period_{p,it}) + \epsilon_{it}$$
+
+where $K = \{$Upgraded, Newly-Installed$\}$
+
+These specifications allow us to identify which types of infrastructure improvements and implementation strategies are most effective at increasing Bike Share utilization.
+For both heterogeneity analyses, we cluster standard errors at the bikeway level to account for serial correlation within bikeways over time.
+
+# Results {#sec-results}
+## Parallel Trends Analysis
+A key assumption of our difference-in-differences design is that treated and control bikeways would have followed parallel trends in the absence of treatment.
+@fig-parallel visualizes this assumption by plotting the seasonally adjusted monthly rides for both treated and control bikeways during the pre-treatment period.
+While the treatment group consistently shows higher monthly average ridership levels than the control group, the trends between the two groups appear to move in parallel before treatment, particularly in the period from month -20 to month -10.
+We exclude the six months immediately preceding treatment from this visual analysis due to possible overlap in construction, avoiding potential anticipation effects.
+
+```{r}
+#| label: fig-parallel
+#| fig-cap: "Parallel Trends Test: Average Monthly Rides in Pre-Treatment Period"
+#| echo: false
+#| warning: false
+#| message: false
+
+parallel_plot <- analysis_data %>%
+  filter(relative_month <= -6)
+
+ggplot(parallel_plot, 
+       aes(x = relative_month, 
+           y = monthly_rides_adj, color = treatment)) +
+  stat_summary(geom = "line", fun = mean, size = 1) +
+  labs(
+    x = "Relative Months (Pre-Treatment)",
+    y = "Seasonally Adjusted Monthly Rides",
+    color = "Treatment Status"
+  )  +
+  scale_color_manual(values = c("TRUE" = "#ff7f0e", "FALSE" = "#1f77b4"),
+                     labels = c("Control", "Treated")) +
+  theme_minimal() +
+  theme(legend.position = "top",
+        panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
+        )
+
+```
+
+The full results of the parallel trends test are provided in @tbl-parallel in the appendix.
+The coefficients show no clear trend and are generally not statistically different from zero in the pre-treatment period, supporting the parallel trends assumption.
+While we observe slightly negative coefficients in months -8 and -7 (significant at the 10% and 5% levels respectively), the magnitudes are small relative to our main treatment effects and may reflect normal variation rather than systematic differences.
+
+## Difference-in-Differences Model Estimates
+@tbl-main-results presents the results from our main difference-in-differences model.
+The model estimates that infrastructure improvements led to a significant increase in Bike Share ridership.
+During the treatment year, we observe a marginally significant increase of 32.4 rides per month (p < 0.10).
+The effect becomes substantially larger in the post-treatment period, with an estimated increase of 133.1 rides per month (p < 0.01), compared to the pre-treatment period.
+
+The high R-squared value of 0.785 indicates that our model explains a substantial portion of the variation in ridership.
+The inclusion of bikeway and time fixed effects helps control for unobserved confounders that might affect ridership patterns.
+
+```{r}
+#| label: tbl-main-results
+#| tbl-cap: "Main Difference-in-Difference Results"
+#| echo: false
+#| warning: false
+#| message: false
+
+main_did <- readRDS(here::here("models/main_did.rds"))
+
+modelsummary(
+  list("Main DiD Model" = main_did),
+  stars = TRUE,
+  gof_map = c("nobs", "r.squared", "adj.r.squared")
+)
+```
 
 ## Robustness Checks
+
 ### Placebo Test Results
-- Present results and discuss implications
+To verify that our results are not driven by spurious correlations, we conduct a placebo test by randomly reassigning treatment status across bikeways.
+As shown in @tbl-placebo, when using this random assignment, we find near-zero effects in both the treatment period (4.0 rides) and post-treatment period (0.1 rides), with neither coefficient being statistically significant.
+These null results, compared to the significant effects found in our main analysis, support the validity of our findings by demonstrating that randomly assigned "fake" treatments do not generate the ridership increases we observe with actual infrastructure improvements.
 
-### Heterogenous Effects
-- Compare effects across bikeway types
-- Include figures or tables
+```{r}
+#| label: tbl-placebo
+#| tbl-cap: "Placebo Test Results - Random Treatment Assignment"
+#| echo: false
+#| warning: false
+#| message: false
 
-# Discussion
-## Model Results Interpretation
+placebo_random_did <- readRDS(here::here("models/placebo_random_did.rds"))
 
-## Policy Implications and Recommendations
+# Randomly assign treatment
+set.seed(778)
 
-## Limitations and Future Research
+placebo_random <- analysis_data %>%
+  # First get a list of unique bikeway IDs
+  group_by(bikeway_id) %>%
+  # Take the first row of each group to get treatment status
+  slice(1) %>%
+  # Randomly shuffle the treatment status across bikeways
+  ungroup() %>%
+  mutate(placebo_treatment = sample(treatment, n(), replace = FALSE)) %>%
+  # Now join this back to the original data
+  select(bikeway_id, placebo_treatment) %>%
+  right_join(analysis_data, by = "bikeway_id")
+
+modelsummary(
+  list("Placebo Test - Random Assignment" = placebo_random_did),
+  stars = TRUE,
+  gof_map = c("nobs", "r.squared", "adj.r.squared")
+)
+
+```
+
+### Heterogeneous Effects
+We examine treatment effect heterogeneity across different types of infrastructure and implementation strategies.
+@tbl-hetero-lane presents the results by bikeway type, with each coefficient representing the change in ridership compared to the pre-treatment period for that specific infrastructure type.
+Protected lanes show the most dramatic impact, generating an additional 381.9 rides per month post-treatment (p < 0.001) – more than twice the effect of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001).
+The substantially larger effect of protected lanes suggests that perceived safety significantly influences riders' willingness to use Bike Share.
+
+```{r}
+#| label: tbl-hetero-lane
+#| tbl-cap: "Heterogeneous Treatment Effects by Bikeway Type"
+#| echo: false
+#| warning: false
+#| message: false
+
+did_hetero_lane <- readRDS(here::here("models/did_hetero_lane_type.rds"))
+
+modelsummary(
+  list("Heterogeneous Effects - Bikeway Type" = did_hetero_lane),
+  stars = TRUE,
+  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  coef_map = c(
+  "period::treatment:treatment_infra::Control" = "Treatment x Control",
+  "period::treatment:treatment_infra::Treated-OnRoad" = "Treatment x Treated-OnRoad",
+  "period::treatment:treatment_infra::Treated-Protected" = "Treatment x Treated-Protected",
+  "period::treatment:treatment_infra::Treated-Shared" = "Treatment x Treated-Shared",
+  "period::post:treatment_infra::Control" = "Post x Control",
+  "period::post:treatment_infra::Treated-OnRoad" = "Post x Treated-OnRoad",
+  "period::post:treatment_infra::Treated-Protected" = "Post x Treated-Protected",
+  "period::post:treatment_infra::Treated-Shared" = "Post x Treated-Shared"
+  ),
+  output = "latex"
+)
+```
+
+@tbl-hetero-treat compares the effectiveness of upgrading existing infrastructure versus installing new bikeways.
+Upgrades to existing infrastructure generate larger and more immediate benefits, showing both a significant treatment-year effect (43.8 additional rides, p < 0.10) and a strong post-treatment effect (152.7 additional rides, p < 0.01).
+In contrast, newly-installed bikeways show no significant effect during the treatment year and a smaller post-treatment increase (115.5 rides, p < 0.10), suggesting that improvements to established cycling routes may be more effective at increasing ridership than creating entirely new routes.
+
+```{r}
+#| label: tbl-hetero-treat
+#| tbl-cap: "Heterogeneous Treatment Effects by Implementation Strategy"
+#| echo: false
+#| warning: false
+#| message: false
+
+did_hetero_treat_type <- readRDS(here::here("models/did_hetero_treat_type.rds"))
+
+modelsummary(
+  list("Heterogeneous Effects - Treatment Type" = did_hetero_treat_type),
+  stars = TRUE,
+  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  coef_map = c(
+    "period::treatment:sub_treatment_type::Upgraded" = "Treatment x Upgraded",
+    "period::treatment:sub_treatment_type::Newly-Installed" = "Treatment x Newly Installed",
+    "period::post:sub_treatment_type::Upgraded" = "Post x Upgraded",
+    "period::post:sub_treatment_type::Newly-Installed" = "Post x Newly Installed"
+  ),
+  output = "latex"
+)
+```
+
+These heterogeneity analyses reveal important nuances in the effectiveness of different infrastructure strategies.
+The particularly strong effect of protected lanes suggests that safety and separation from vehicle traffic may be key factors in encouraging Bike Share usage.
+Additionally, the stronger effects of upgrades compared to new installations may indicate the importance of improving infrastructure in areas where cycling activity already exists.
+
+# Discussion {#sec-discussion}
+Our difference-in-differences analysis of Toronto's cycling infrastructure changes between 2019-2021 reveals nuanced insights about how different types of cycling infrastructure improvements influence Bike Share utilization.
+The results demonstrate that while infrastructure investments broadly increase cycling adoption, the magnitude and timing of these effects vary substantially by implementation approach and infrastructure type.
+
+## Key Empirical Findings and Their Implications
+The evolution of treatment effects over time tells an important story about behavior change and infrastructure adoption.
+During the treatment year, we observe a modest but marginally significant increase of 32.4 rides per month (p < 0.10).
+This relatively small initial effect likely reflects several factors.
+First, infrastructure improvements often cause temporary disruption during construction, potentially suppressing immediate usage of bike share in the surrounding areas.
+Second, the treatment year impacts may be diluted because we only observe the year of changes, not precise implementation dates.
+Most importantly, the minor initial increase suggests that behavioral adaptation to new infrastructure takes time as users adjust their travel patterns and routine to new routes.
+
+The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – reveals the true potential of infrastructure improvements once fully integrated into the transportation network.
+This impact from pre-treatment to post-treatment periods aligns with research from the University of Colorado Denver showing that "building safe facilities for cyclists is one of the biggest factors in road safety for everyone" [@MARSHALL2019100539].
+The magnitude of this change suggests that infrastructure improvements don't just make existing cyclists' trips safer, they also alter the decisions behind mode of transportation for many residents.
+
+To put these effects in perspective, our treatment estimate of 133.1 additional monthly rides per bikeway translates to approximately 24,757 additional monthly rides across all 186 treated segments, or nearly 300,000 additional rides annually.
+This substantial increase is reflected in the raw data (see @tbl-summary-statistics in Appendix @sec-additional-figures) - treated segments averaged 331.92 monthly rides post-treatment compared to 209.03 pre-treatment, while control segments saw a more modest increase from 85.59 to 107.77 monthly rides over the same period.
+The magnitude of this effect is particularly notable given that treated segments represent only about 16% of Toronto's total bikeways (186 out of 1,191 total segments) but generated significant ridership gains, suggesting that targeted infrastructure improvements can have outsized impacts on system utilization.
+
+The dramatic difference in baseline ridership between treatment and control groups (209.03 versus 85.59 average monthly rides pre-treatment, as shown in @tbl-summary-statistics) also provides important context for interpreting our heterogeneous effects.
+This pattern suggests that the city prioritized improvements on already-popular routes, which may partially explain why upgrades to existing infrastructure showed more immediate effects than new installations.
+However, the fact that treated segments still demonstrated substantial gains in ridership despite starting from a higher baseline suggests that infrastructure improvements can drive meaningful increases in cycling activity even on already-established routes.
+
+This substantial increase in post-treatment ridership may also be partially driven by strategic expansion of the Bike Share network itself.
+Following infrastructure improvements, Bike Share Toronto can coordinate with the city to place new stations near cycling infrastructure.
+This complementary expansion of both infrastructure and Bike Share access could create a multiplier effect, where improved infrastructure attracts more stations, which in turn enables more riders to access the enhanced bike lanes.
+While this interaction between infrastructure and system expansion complicates our causal interpretation, it reflects the real-world synergy between cycling infrastructure and bike share system development.
+
+### Heterogenous Effect Findings
+Our heterogeneous effects analysis reveals crucial insights about which interventions most effectively promote cycling.
+Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment (p < 0.001) – more than twice the impact of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001).
+This substantial difference in effectiveness likely stems from:
+
+- **Safety perceptions**: The physical separation from traffic provided by protected lanes addresses the concerns of what @nacto2016 research identifies as the "interested but concerned" majority of potential cyclists.
+- **User confidence**: Protected infrastructure may encourage less experienced cyclists to try Bike Share, expanding the user base beyond confident urban cyclists.
+- **Urban density patterns**: Protected lanes are more commonly implemented in denser, high-traffic areas like downtown Toronto, where higher population density and concentration of destinations naturally generate more cycling trips.
+This geographic distribution may partially explain their stronger treatment effects compared to other infrastructure types.
+
+The comparison between upgrades and new installations provides equally important insights for implementation strategy.
+Upgrades to existing infrastructure generate more monthly ridership (152.7 adjusted rides), compared to newly-installing bike lanes (115.5 adjusted rides) post-treatment.
+Additionally, during the treatment period, upgrades to existing infrastructure show immediate effect to ridership (43.8 rides, p < 0.10), whereas new installations show no significant treatment-year effect.
+This pattern suggests that improving existing cycling routes may be more effective than expanding into new areas, particularly for short-term ridership gains.
+This finding aligns with Toronto's strategic focus during the same period as our study, on upgrading existing infrastructure to "higher-order" facilities, as outlined in the 2019-2021 Cycling Network Plan Update [@toronto_council_agenda].
+
+## Infrastucture Development and Policy Implications
+Our findings have substantial implications for cycling infrastructure development strategies.
+The strong performance of protected lanes, combined with the effectiveness of upgrades, suggests a coupled approach:
+
+- Systematically upgrade existing high-traffic routes to protected infrastructure where feasible
+- Strategically expand the network through new protected lanes in high-potential corridors
+
+This approach is particularly relevant given Toronto's commitment to add 100 kilometers of new cycling infrastructure between 2025 and 2027 [@cyclingNetworkPlan].
+The significant post-treatment effects we observe – especially for protected lanes – suggest these investments could substantially increase Bike Share ridership if properly implemented.
+These findings and implications align with historical actions, as in 2020, the City of Toronto worked to expand Bike Share coverage to 20 of the city's 25 wards (Toronto Parking Authority 2020), further suggesting opportunities for coordinated deployment of infrastructure improvements and system expansion.
+
+The timing of effects also has important implications for policy evaluation.
+The delay between implementation and peak impact suggests that infrastructure evaluations should maintain longer time horizons.
+Short-term metrics may significantly underestimate the full benefits of cycling investments, particularly for new installations which show no significant treatment-year effects but meaningful long-term impacts.
+This delayed-benefit pattern aligns with experiences in other cities; for instance, Portland saw a 75% reduction in road fatality rates over a 20-year period of sustained infrastructure investment, highlighting the importance of maintaining long-term commitment to cycling infrastructure development [@MARSHALL2019100539].
+
+## Methodological Strengths and Limitations
+Our difference-in-differences design, supported by robust parallel trends in the pre-treatment period and null results in placebo tests, provides strong causal evidence for the effectiveness of cycling infrastructure.
+The robustness of our results is particularly noteworthy given Toronto's challenging research context, where multiple infrastructure projects and transportation initiatives often overlap temporally and spatially.
+However, several limitations merit discussion:
+
+- **Temporal precision**: Our inability to observe exact implementation dates within treatment years likely diminishes treatment-year effects and may understate immediate post-implementation impacts.
+- **Spatial relationships**: While we control for station proximity to bikeways, we cannot observe actual route choices.
+This may lead us to underestimate infrastructure usage if riders access lanes from more distant starting points, or even overestimate usage if riders don't use bike lanes, despite starting near one.
+This limitation is particularly significant in downtown areas where the density of both bike lanes and stations makes it difficult to isolate specific infrastructure effects.
+- **Station placement effects**: Our results may partially reflect strategic placement of new Bike Share stations near improved infrastructure rather than purely infrastructure-driven changes in behavior.
+While this interaction makes it harder to isolate the direct effect of infrastructure improvements, it captures the realistic policy dynamic where bike share systems expand in tandem with cycling infrastructure.
+
+## Future Research Directions
+Several promising avenues for future research emerge from these findings:
+
+- **User surveys**: Qualitative research could help explain the mechanisms behind the stronger effects of protected lanes and how infrastructure influences mode choice decisions.
+Such research could particularly focus on understanding the "interested but concerned" demographic identified by @nacto2016 as crucial for expanding cycling adoption.
+- **Network effects**: Future studies could examine how the benefits of infrastructure improvements spill over onto connecting routes and affect system-wide ridership patterns.
+This could help inform Toronto's goal of creating a more connected cycling network, as outlined in the city's 2025-2027 implementation program [@cyclingNetworkPlan].
+- **Neighborhood-level analysis**: Future research could examine how infrastructure effects vary across Toronto's diverse neighborhoods.
+Such analysis would be particularly valuable as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact.
+This geographic heterogeneity analysis could also inform equity considerations in infrastructure deployment, ensuring benefits are distributed across the city's diverse communities.
+- **Integration with public transit**: Studies could investigate how cycling infrastructure near transit hubs affects multi-modal trip patterns, particularly relevant given Toronto's efforts to integrate Bike Share with the TTC system [@ttcIntegration].
+
+This research contributes to our understanding of how cycling infrastructure influences urban mobility patterns.
+The clear relationship between infrastructure quality and ridership, particularly the strong effect of protected lanes, provides evidence-based guidance for cities working to encourage sustainable transportation.
+As Toronto continues expanding its cycling network, these insights can help inform investment priorities and implementation strategies to maximize the impact of infrastructure improvements.
 
@@
 # Appendix {.unnumbered}
-# Idealized Survey Methodology
+# Idealized Survey Methodology {#sec-ideal-survey}
 ## Survey Objectives
@@
 
-3.
+3. **Attention Check**: "Select the option 'Green'
+
+  > - Red
+  > - Elephant
+  > - Green
+  > - Circle
+  > - Rain
+
+4.
 **Visual Choice Experiment** (Show side-by-side images of the same bike lane before and after upgrades) : "If you needed to travel along this route, which version would make you more likely to use Bike Share?
@@
 
-4. **Behavioral Scenarios**: "Imagine a Bike Share station gets installed near your workplace, and the city adds protected bike lanes on your route to work. What would you do?
+5. **Behavioral Scenarios**: "Imagine a Bike Share station gets installed near your workplace, and the city adds protected bike lanes on your route to work. What would you do?
 
@@
 
-# Additional Tables & Figures
+\newpage
+# Data Processing Methodology {#sec-data-processing}
+This appendix section provides a detailed overview of our data processing methodology, explaining how we construct the final analysis dataset from raw Bike Share and cycling infrastructure data.
 
-# Data Processing
 ## Spatial Filtering
-## Treatment and Control Group Construction
+We begin by identifying bike stations near cycling infrastructure:
 
-# Model Details and Diagnostics
+1. Calculate distances between each Bike Share station and nearby bikeways
+2. Retain only stations within 100 meters of a bikeway
+3. Tag each retained station with its nearest bikeway ID
+
+This spatial filtering reduces our dataset from 21.8 million total rides to 13.8 million rides starting near bikeways, creating a focused sample of trips most likely to utilize cycling infrastructure.
+
+## Monthly Ridership Aggregation and Adjustment
+To prepare for our difference-in-differences analysis, we:
+
+1. Aggregate rides by bikeway and month
+2. Create seasonal adjustment factors based on monthly patterns
+3. Apply these adjustments to normalize seasonal variation in ridership
+
+## Treatment and Control Groups
+**Treatment Group**
+
+- Identify bikeways upgraded or constructed in 2019-2021
+- Create 60-month windows for each:
+  * 24 months pre-treatment
+  * 12 months treatment year
+  * 24 months post-treatment
+
+**Control Group**
+
+- Select bikeways unchanged since pre-2017
+- Randomly assign pseudo-treatment years (2019-2021)
+- Create matching 60-month observation windows
+
+This approach yields our final analysis dataset of approximately 8 million rides, capturing ridership patterns around infrastructure changes while controlling for broader trends through our comparison group.
+
+## Key Methodological Decisions
+
+1. **100-meter Threshold**: Balances capturing relevant rides while maintaining confidence in infrastructure usage
+2. **60-month Window**: Provides sufficient pre/post data while maintaining balanced observations
+3. **Seasonal Adjustment**: Addresses Toronto's strong cycling seasonality to better isolate infrastructure effects
+
+These processing steps create a dataset structured to identify the causal impact of infrastructure changes on ridership patterns.
+
+# Additional Tables & Figures {#sec-additional-figures}
+```{r}
+#| label: tbl-sample-data
+#| tbl-cap: "Sample of processed dataset for DiD Model"
+#| echo: false
+#| warning: false
+#| message: false
+
+set.seed(101)
+
+sample_data <- analysis_data %>% sample_n(5) %>% 
+  as_tibble() %>% arrange(year_month) %>% t()
+
+# Generate the table with kable and kableExtra
+kable(sample_data, format.args = list(big.mark = ",")) %>%
+  kable_styling(font_size = 8) %>%
+  column_spec(1, width = "3cm", bold = TRUE) %>%
+  column_spec(2:ncol(sample_data), width = "2cm")
+```
+
+```{r}
+#| label: fig-bikeway_type_constructed_graph
+#| fig-cap: "Breakdown of Bikeway Types by Installation Year (2002-2023)"
+#| echo: false
+#| warning: false
+#| message: false
+
+bikeWay_data <- read_parquet(here::here("data/02-analysis_data/bikeway_data.parquet"))
+
+bikeway_colors <- c("Shared Roadways" = "#e15759", 
+                    "On-Road Lanes" = "#76b7b2",
+                    "Protected Lanes" = "#2ca02c")
+
+bikeWay_type_summary <- bikeWay_data %>%
+  filter(INSTALLED >= 2002 & INSTALLED <= 2023) %>%
+  group_by(INSTALLED, INFRA_HIGHORDER) %>%
+  summarise(count = n())
+
+ggplot(bikeWay_type_summary, aes(x = as.factor(INSTALLED), y = count, fill = INFRA_HIGHORDER)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = bikeway_colors) +
+  labs(
+    x = "Year Installed",
+    y = "Count of Bikeways",
+    fill = "Bikeway Types:"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    panel.grid.major = element_line(color = "gray85", linewidth = 0.35),
+    legend.position = "top"
+  ) +
+  annotate(
+    "rect",
+    xmin = which(levels(as.factor(bikeWay_type_summary$INSTALLED)) == "2019") - 0.5,
+    xmax = which(levels(as.factor(bikeWay_type_summary$INSTALLED)) == "2021") + 0.5,
+    ymin = -Inf,
+    ymax = Inf,
+    alpha = 0.2,
+    fill = "gray20"
+  )
+```
+
+```{r}
+#| label: fig-bikeway_type_upgraded_graph
+#| fig-cap: "Breakdown of Bikeway Types by Upgraded Year (2002-2023)"
+#| echo: false
+#| warning: false
+#| message: false
+
+bikeWay_type_summary <- bikeWay_data %>%
+  filter(UPGRADED >= 2002 & UPGRADED <= 2023) %>%
+  group_by(UPGRADED, INFRA_HIGHORDER) %>%
+  summarise(count = n()) %>%
+  drop_na(UPGRADED)
+
+ggplot(bikeWay_type_summary, aes(x = as.factor(UPGRADED), y = count, fill = INFRA_HIGHORDER)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = bikeway_colors) +
+  labs(
+    x = "Year Upgraded",
+    y = "Count of Bikeways",
+    fill = "Bikeway Types:"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    panel.grid.major = element_line(color = "gray85", linewidth=0.35),
+    legend.position = "top"
+  ) +
+  annotate(
+    "rect",
+    xmin = which(levels(as.factor(bikeWay_type_summary$UPGRADED)) == "2019") - 0.5,
+    xmax = which(levels(as.factor(bikeWay_type_summary$UPGRADED)) == "2021") + 0.5,
+    ymin = -Inf,
+    ymax = Inf,
+    alpha = 0.2,
+    fill = "gray20"
+  )
+```
+
+```{r}
+#| label: tbl-parallel
+#| tbl-cap: "Parallel Trends Test Results - Monthly Treatment Effects"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare pre-treatment data for parallel trends test
+pre_treatment_data <- analysis_data %>%
+  filter(period == "pre" & relative_month < -5) %>%
+  mutate(relative_month_treatment_interaction = interaction(
+    relative_month, treatment, drop = TRUE
+  ))
+
+parallel_trends_test <- readRDS(here::here("models/parallel_trends_test.rds"))
+
+modelsummary(
+  list("Coefficient (std.error)" = parallel_trends_test),
+  stars = TRUE,
+  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  data = pre_treatment_data,
+  estimate = "{estimate}{stars} ({std.error})", 
+  statistic = NULL
+)
+```
+
+\newpage
+## Summary Statistics
+@tbl-summary-statistics is a summary statistics table comparing the control and treatment groups ridership and bikeway details.
+The Pre-period is 24 relative months before the treatment period, and Post-period is 24 relative months after the treatment period, while the Treatment-period spans 12 months.
+
+```{r}
+#| label: tbl-summary-statistics
+#| tbl-cap: "Summary Statistics between treatment and control groups"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summary statistics for total rides, mean monthly ridership, and number of bikeways
+summary_stats <- analysis_data %>%
+  group_by(treatment, period) %>%
+  summarise(
+    total_rides = sum(monthly_rides, na.rm = TRUE),
+    mean_rides_adj = mean(monthly_rides_adj, na.rm = TRUE),
+    n_bikeways = n_distinct(bikeway_id),
+    .groups = "drop"
+  ) %>%
+  pivot_wider(
+    names_from = period,
+    values_from = c(total_rides, mean_rides_adj, n_bikeways),
+    names_glue = "{.value}_{period}"
+  )
+
+# Distribution of infrastructure types
+infra_dist <- analysis_data %>%
+  filter(period %in% c("pre", "post")) %>% # Focus on pre/post periods
+  group_by(treatment) %>%
+  summarise(
+    protected = mean(bikeway_type == "Protected Lanes", na.rm = TRUE),
+    onroad = mean(bikeway_type == "On-Road Lanes", na.rm = TRUE),
+    shared = mean(bikeway_type == "Shared Roadways", na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Combine statistics and infrastructure distributions
+final_table <- summary_stats %>%
+  left_join(infra_dist, by = "treatment") %>%
+  pivot_longer(
+    cols = -treatment,
+    names_to = "metric",
+    values_to = "value"
+  ) %>%
+  filter(!str_detect(metric, "n_bikeways_") | metric == "n_bikeways_post") %>% # Keep only `n_bikeways_post`
+  pivot_wider(
+    names_from = treatment,
+    values_from = value,
+    names_prefix = "Group_"
+  ) %>%
+  rename(
+    `Metric` = metric,
+    `Control` = Group_FALSE,
+    `Treatment` = Group_TRUE
+  )
+
+# Clean up Metric Names
+final_table <- final_table %>%
+  mutate(
+    Metric = case_when(
+      Metric == "total_rides_post" ~ "Total Rides (Post)",
+      Metric == "total_rides_pre" ~ "Total Rides (Pre)",
+      Metric == "total_rides_treatment" ~ "Total Rides (Treatment)",
+      Metric == "mean_rides_adj_post" ~ "Mean Rides Adjusted (Post)",
+      Metric == "mean_rides_adj_pre" ~ "Mean Rides Adjusted (Pre)",
+      Metric == "mean_rides_adj_treatment" ~ "Mean Rides Adjusted (Treatment)",
+      Metric == "n_bikeways_post" ~ "Number of Bikeways",
+      Metric == "protected" ~ "Protected Lanes",
+      Metric == "onroad" ~ "On-Road Lanes",
+      Metric == "shared" ~ "Shared Roadways",
+      TRUE ~ Metric
+    )
+  )
+
+final_table %>%
+  mutate(across(c(Control, Treatment), ~ round(.x, 2))) %>%  # Round to 0 decimal places
+  kable(
+    format = "pipe",
+    col.names = c("Metric", "Control Group", "Treatment Group"),
+    format.args = list(big.mark = ",")  # Add commas for large numbers
+  )
+
+```
 

### 2024-12-02 scripts/03-clean_data.R

@@ -6,7 +6,7 @@
 # Pre-requisites: 
 # - 02-download_data.R must have been run first
 # - The required libraries listed below must be installed:
-#     - readr, tidyverse, fs, lubridate
+#     - tidyverse, dplyr, readr, lubridate, fs, sf, units, arrow
 
 #### Workspace setup ####
 library(tidyverse)
@@ -17,17 +17,6 @@ library(fs)
 library(sf)
 library(units)
 library(arrow)
-
-############################# Manage Directories #############################
-# # Create directory to store cleaned data
-# analysis_data_dir <- "data/02-analysis_data"
-# 
-# # Check if the directory exists, if it does, delete it
-# if (dir_exists(analysis_data_dir)) {
-#   dir_delete(analysis_data_dir)  # Delete the folder and its contents
-# }
-# 
-# dir_create(analysis_data_dir)
 
 
 ########################### Cleaning Bike Way Data ############################
@@ -192,7 +181,6 @@ valid_stations <- final_stations$station_id
 filtered_rides <- combined_rides %>%
   filter(
     start_station_id %in% valid_stations
-    # , end_station_id %in% valid_stations # If filtering for both stations near bikeway
   ) %>%
   # Join start station details
   left_join(
@@ -210,37 +198,32 @@ filtered_rides <- combined_rides %>%
              end_nearest_bikeway = nearest_bikeway_id),
     by = c("end_station_id" = "station_id")
   )
-  # filter(start_nearest_bikeway == end_nearest_bikeway) # If filtering for both stations near same bikeway
-
-## Save filtered rides as CSV
-# write_csv(filtered_rides,
-#           "data/02-analysis_data/filtered_ridership.csv")
 
 ################ Create Final Dataset for Diff-in-Diff model ##################
-# Step 1: Identify treatment and control bikeways
+###### Identify treatment and control bikeways
 bikeway_status <- bike_lanes %>%
   mutate(
     # Determine treatment year (either UPGRADED or INSTALLED if no upgrade)
-    upgrade_installed_year = case_when(UPGRADED > 0 ~ UPGRADED
+    construct_year = case_when(UPGRADED > 0 ~ UPGRADED
                                        , TRUE ~ INSTALLED),
     
     # Flag if bikeway was treated in our window of interest (2019-2021)
[…]

### 2024-12-02 scripts/04-test_analysis_data.R

@@ -1,69 +1,159 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Testing cleaned bike share and bike lane data
+# Author: Steven Li
+# Date: 1 December 2024
+# Contact: stevency.li@mail.utoronto.ca
+# Pre-requisites: 
+# - 03-clean_data.R must have been run first
+# - The required packages must be installed:
+#     - tidyverse, testthat, arrow
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Read in the cleaned data
+bike_lanes <- read_parquet("data/02-analysis_data/bikeway_data.parquet")
+final_df <- read_parquet("data/02-analysis_data/final_df.parquet")
 
-
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+######################## Test Bike Lanes Data Structure ######################
+test_that("Bike lanes data has correct structure", {
+  # Test presence of required columns
+  expected_columns <- c("OBJECTID", "INSTALLED", "UPGRADED", "INFRA_HIGHORDER")
+  expect_true(all(expected_columns %in% colnames(bike_lanes)))
+  
+  # Test data types
+  expect_true(is.numeric(bike_lanes$OBJECTID))
+  expect_true(is.numeric(bike_lanes$INSTALLED))
+  expect_true(is.numeric(bike_lanes$UPGRADED))
+  expect_true(is.character(bike_lanes$INFRA_HIGHORDER))
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+test_that("Bike lane types are valid", {
+  valid_types <- c("Protected Lanes", "On-Road Lanes", "Shared Roadways")
+  
+  # Test that all types are valid
+  expect_true(all(bike_lanes$INFRA_HIGHORDER %in% valid_types))
[…]

### 2024-12-02 scripts/05-model_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
 # Purpose: Run Difference-in-Difference model and robustness tests
 # Author: Steven Li
-# Date: 28 November 2024
+# Date: 1 December 2024
 # Contact: stevency.li@mail.utoronto.ca
 # Pre-requisites: 
-# - '02-download_data.R' and '03-clean_data.R' must have been run first
+# - '02-download_data.R', '03-clean_data.R' and '04-test_analysis_data.R' must have been run first
 # - The required libraries listed below must be installed:
 #     - tidyverse, fixest, modelsummary, and arrow
 
@@ -20,13 +20,15 @@ library(arrow)
 analysis_data <- read_parquet("data/02-analysis_data/final_df.parquet")
 
 ################################ Basic DiD Model ############################
+# Create main DiD model, with pre-treatment period as reference
 main_did <- feols(
-  bikeway_monthly_rides_adjusted ~ 
+  monthly_rides_adj ~ 
     i(period, treatment, ref = "pre") | 
     bikeway_id + relative_month,            
   data = analysis_data
 )
 
+# Save model
 saveRDS(main_did, file="models/main_did.rds")
 
 # # Outut model stats
@@ -41,25 +43,26 @@ saveRDS(main_did, file="models/main_did.rds")
 set.seed(778)
 
 placebo_random <- analysis_data %>%
-  # First get a list of unique bikeway IDs
+  # Get a list of unique bikeway IDs
   group_by(bikeway_id) %>%
   # Take the first row of each group to get treatment status
   slice(1) %>%
   # Randomly shuffle the treatment status across bikeways
   ungroup() %>%
   mutate(placebo_treatment = sample(treatment, n(), replace = FALSE)) %>%
-  # Now join this back to the original data
+  # Join back to the original data
   select(bikeway_id, placebo_treatment) %>%
   right_join(analysis_data, by = "bikeway_id")
 
 # Run the placebo DiD
 placebo_random_did <- feols(
-  bikeway_monthly_rides_adjusted ~ 
+  monthly_rides_adj ~ 
     i(period, placebo_treatment, ref = "pre") | 
     bikeway_id + relative_month,
   data = placebo_random
 )
 
+# Save model
 saveRDS(placebo_random_did, file="models/placebo_random_did.rds")
 
 # # Outut model stats
[…]

### 2024-12-02 scripts/02-download_data.R
commenting out unneeded scripts

@@ -32,11 +32,6 @@ library(httr)
 ############################# Manage Directories #############################
 raw_data_dir <- "data/01-raw_data"
 
-# Check if the directory exists, if it does, delete it
-if (dir_exists(raw_data_dir)) {
-  dir_delete(raw_data_dir)  # Delete the folder and its contents
-}
-
 # Recreate the directories after deletion
 bikeshare_file_path <- file.path(raw_data_dir, "01-raw_bikeshare_data")
 dir_create(bikeshare_file_path)
@@ -80,49 +75,49 @@ for(i in 1:nrow(datastore_resources)) {
   }
 }
 
-########################## Bike Share Stations data ##########################
-# get package details
-package <- show_package("2b44db0d-eea9-442d-b038-79335368ad5a")
-
-# get all resources for this package
-resources <- list_package_resources("2b44db0d-eea9-442d-b038-79335368ad5a")
-
-# filter relevant file needed (bike-share-json)
-datastore_resources <- filter(resources, name == 'bike-share-json')
-
-# Get metadata (contains relevant URL to json)
-gbfs_data <- get_resource(datastore_resources$id)
-
-# Extract the station_information URL
-station_url <- gbfs_data$data$en$feeds$url[3]  # Third URL is station_information
-
-# Use httr to get the station data
-response <- GET(station_url)
-station_data <- fromJSON(rawToChar(response$content))
-
-# Convert to a tibble
-stations_df <- station_data$data$stations %>%
-  as_tibble() %>%
-  select(station_id, name, lat, lon)
-
-#### Save data ####
-file_path <- file.path(bikestation_file_path, "bike_station_data.csv")
-write_csv(stations_df, file_path)
-
-##################### Cycling Network (Bike ways) data ########################
-# get package details
-package <- show_package("cycling-network")
-
-# get all resources for this package
-resources <- list_package_resources("cycling-network")
-
-# identify datastore resources
-datastore_resources <- filter(resources, 
-                              name == 'cycling-network - 4326.geojson')
-
-# load datastore resource
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
-
-#### Save data ####
-file_path <- file.path(bikeway_file_path, "bikeway_data.geojson")
-st_write(data, file_path, driver = "GeoJSON")
+# ########################## Bike Share Stations data ##########################
+# # get package details
+# package <- show_package("2b44db0d-eea9-442d-b038-79335368ad5a")
[…]

### 2024-12-02 scripts/03-clean_data.R
fixed error in cleaning script

@@ -45,41 +45,41 @@ bike_lanes <- bike_lanes %>%
 ### We need geometry data for further data processing. So geometry column will
 ### be dropped at the end, so bike_lanes can be saved as parquet file.
 
-# ################# Add Station ID info to 2017 Q3-Q4 rides  ####################
-# # 2017 Q3 and Q4 are missing station ID data. Will join with bike station info to get ID's
-# bike_station <- read_csv(
-#   "data/01-raw_data/02-raw_bikestation_data/bike_station_data.csv")
-#
-# df2017 <- read_csv(
-#   "data/01-raw_data/01-raw_bikeshare_data/Bikeshare Ridership (2017 Q3).csv")
-#
-# # Left join twice by station name for starting and ending station ID
-# new_temp <- df2017 %>%
-#   left_join(bike_station, by=c("from_station_name" = "name")) %>%
-#   rename(from_station_id = station_id) %>%
-#   left_join(bike_station, by=c("to_station_name" = "name")) %>%
-#   rename(to_station_id = station_id) %>%
-#   select(trip_id, trip_start_time, from_station_id, from_station_name,
-#          to_station_id, to_station_name) %>%
-#   drop_na() # filter rows with non-matching station IDs
-#
-# write_csv(new_temp, "data/01-raw_data/01-raw_bikeshare_data/Bikeshare Ridership (2017 Q3).csv")
-#
-# ### Do the same steps again fr Q4 2017
-# df2017 <- read_csv(
-#   "data/01-raw_data/01-raw_bikeshare_data/Bikeshare Ridership (2017 Q4).csv")
-#
-# # Left join twice by station name for starting and ending station ID
-# new_temp <- df2017 %>%
-#   left_join(bike_station, by=c("from_station_name" = "name")) %>%
-#   rename(from_station_id = station_id) %>%
-#   left_join(bike_station, by=c("to_station_name" = "name")) %>%
-#   rename(to_station_id = station_id) %>%
-#   select(trip_id, trip_start_time, from_station_id, from_station_name,
-#          to_station_id, to_station_name) %>%
-#   drop_na() # filter rows with non-matching station IDs
-#
-# write_csv(new_temp, "data/01-raw_data/01-raw_bikeshare_data/Bikeshare Ridership (2017 Q4).csv")
+################# Add Station ID info to 2017 Q3-Q4 rides  ####################
[…]

### 2024-12-02 other/datasheet/datasheet.qmd
rename datasheet

---
title: "Datasheet for Toronto Cycling Infrastructure Analysis Dataset"
author: 
  - Steven Li
thanks: "Code and data are available at: [https://github.com/stevenli-uoft/Toronto_BikeShare_Causality](https://github.com/stevenli-uoft/Toronto_BikeShare_Causality)."
date: today
date-format: "D MMMM YYYY"
abstract: "This datasheet documents the creation and composition of a dataset analyzing the causal impact of cycling infrastructure improvements on Bike Share Toronto ridership between 2017-2023. Following the standardized format proposed by @gebru2021datasheets, this datasheet details the processing of over 7.4 million bike share rides across 1,191 bikeways, focusing specifically on infrastructure changes during 2019-2021. The dataset combines and cleans records from Toronto's Open Data Portal, including bike share trips, station locations, and cycling infrastructure details, to create a structured dataset for difference-in-differences analysis. The final dataset contains monthly ridership metrics, infrastructure classifications, and temporal indicators designed to evaluate how protected lanes, on-road lanes, and shared roadways influence cycling activity."
format: pdf
number-sections: true
bibliography: references_datasheet.bib
---

All of the following questions are extracted from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the causal effect of cycling infrastructure improvements on Bike Share Toronto ridership between 2017-2023, specifically examining how different types of bikeways influence system usage.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Created by Steven Li at the University of Toronto.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No external funding received.
4. *Any other comments?*
    - No additional comments.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each row represents a bikeway-month observation tracking ridership metrics 24-months  before and after infrastructure changes.
2. *How many instances are there in total (of each type, if appropriate)?*
	- 71,460 observations tracking 1,191 bikeways over 60 months each.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- Dataset includes 1,191 of the roughly 1,500 bikeways in Toronto with associated Bike Share stations within 100m, and Bike Share ridership from 2017-2023.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance contains:

    > - bikeway_id: Unique identifier for each bikeway segment
    > - construct_year: Year when the bikeway was either upgraded or newly installed
    > - period: One of three periods (pre/treatment/post) relative to infrastructure changes
    > - treatment: Boolean indicating if bikeway received improvements during 2019-2021
    > - year_month: Time variable tracking month and year
    > - calendar_year: Year of observation
    > - relative_month: Months relative to treatment (-24 to +24)
    > - bikeway_type: Infrastructure classification (Protected Lanes: Cycle tracks, multi-use trails, bi-directional cycle tracks. On-Road Lanes: Painted bike lanes. Shared Roadways: Sharrows, signed routes, park roads)
    > - sub_treatment_type: Classification of infrastructure change (Upgraded: Existing infrastructure improved. Newly-Installed: New infrastructure added. No Treatment: Control group)
    > - monthly_rides: Raw count of bike share trips starting near bikeway
    > - monthly_rides_adj: Seasonally adjusted ride count
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Treatment status (TRUE/FALSE) indicates whether bikeway received infrastructure improvements during 2019-2021.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- No missing values in final analysis dataset after cleaning.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No explicit relationships between instances.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No recommended splits for analysis dataset.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
  - Spatial precision:

    > - Only captures rides starting within 100m of bikeways
    > - May miss riders who use bikeways but start further away
    > - Some riders within 100m may not use the adjacent bikeway

  - Temporal uncertainty:

    > - Treatment year only known to annual precision
    > - Construction/upgrade timing within year unknown
    > - Seasonal patterns affect ridership (addressed through adjustment)

  - Treatment classification:

    > - Control group randomly assigned to pseudo-treatment years
    > - Some bikeways may have received minor unreported improvements

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Self-contained parquet file.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No confidential information.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No sensitive content.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Bikeways categorized by:

    > - Infrastructure type (Protected Lanes, On-Road Lanes, Shared Roadways)
    > - Treatment type (Upgraded, Newly-Installed, No Treatment)

    - Sub-treatment categorized by:

    > - Newly-Installed: New installation of a bikeway
    > - Upgraded: Upgrades to an existing bikeway
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No individual identification possible.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No sensitive data included.
16. *Any other comments?*
	- No Additional Comments

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
  - Data sourced from City of Toronto Open Data Portal [@torontoOpenData]:

    > - Bike Share ridership data
    > - Bike Share station locations
    > - Cycling Network (Bikeways) data

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

    1. Initial Data Download:

          - Used `opendatatoronto` package [@torontoOpenData] in R [@RCoreTeam2023] to access raw data
          - Downloaded separate datasets for:

              - Bike Share ridership (2017-2023)
              - Bike Share station locations
              - Cycling Network (Bikeways)

    2. Spatial Processing:

          - Using the sf [@sf] package for spatial processing
          - Converted station locations to spatial points using sf package
          - Transformed coordinates to UTM zone 17N for accurate distance calculation
          - Calculated distances between stations and bikeways
          - Filtered to stations within 100m of bikeways

    3. Temporal Processing:

          - Standardized timestamps across datasets
          - Created relative time variables for diff-in-diff analysis
          - Generated seasonal indices for ridership adjustment

    4. Data Integration:

          - Joined ridership data with nearby bikeways
          - Created treatment/control groups based on 2019-2021 changes
          - Generated monthly aggregates of ridership

All processing was validated through unit tests and intermediate data checks, using R packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], and `sf` [@sf] for spatial analysis.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Full population of relevant bikeways included.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data collection automated through R scripts.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Original data spans 2017-2023
	- Data processing completed December 2024
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No ethical review required for public data.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- All data obtained through Toronto Open Data Portal.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Not applicable for public administrative data.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not applicable for public administrative data.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Not applicable for public administrative data.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not applicable for public administrative data.
12. *Any other comments?*
	- No additional comments.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*

    1. Station Data Cleaning:

          - Standardized coordinate systems
          - Removed stations with invalid coordinates
          - Generated unique station-bikeway pairs within 100m

    2. Ridership Data Standardization:

          - Unified column names across years
          - Converted timestamps to consistent format
          - Filtered invalid trip records
          - Added missing station IDs for 2017 Q3-Q4

    3. Bikeway Classification:

          - Categorized infrastructure types from detailed classes
          - Created treatment indicators based on construction/upgrade years
          - Assigned control bikeways to balanced pseudo-treatment years

    4. Temporal Adjustments:

          - Created seasonal indices based on monthly patterns
          - Adjusted ridership for seasonality
          - Generated relative time variables (-24 to +24 months)

    5. Final Integration:

          - Joined cleaned datasets
          - Generated monthly aggregates
          - Created final analysis variables

All cleaning steps are documented in the `03-clean_data.R` script.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Raw data preserved in data/01-raw_data directory
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- All processing done in R
4. *Any other comments?*
	- None

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Analysis of causal impact of cycling infrastructure on bikeshare usage
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- GitHub: https://github.com/stevenli-uoft/Toronto_BikeShare_Causality
3. *What (other) tasks could the dataset be used for?*
  - Evaluating infrastructure investment impacts
  - Planning future cycling network expansion
  - Studying seasonal cycling patterns
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Results should be interpreted considering 100m station proximity threshold
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- Should not be used to assess individual rider behavior
6. *Any other comments?*
	- None

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Publicly available via GitHub
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Parquet file on GitHub
  - No DOI assigned
3. *When will the dataset be distributed?*
	- Available December 2024
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No, MIT License
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No additional restrictions or controls
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No additional restrictions or controls
7. *Any other comments?*
	- None

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Maintained by Steven Li
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Email: stevency.li@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No known errors
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No planned updates
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Not applicable (public data)
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Single version maintained
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Contributions welcome via GitHub pull requests
8. *Any other comments?*
	- None

\newpage

# References

### 2024-12-02 paper/paper.qmd
final paper edits

@@
 ---
-title: "Beyond Infrastructure: Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
-subtitle: "A Difference-in-Differences Analysis of Toronto's Cycling Network, 2017-2023"
+title: "Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
+subtitle: "A Difference-in-Differences Analysis Shows Protected Bike Lanes Triple Bike Share Ridership in Toronto, 2017-2023"
 author: 
@@
 # Introduction
-Cycling infrastructure plays a crucial role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
-While protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
-This knowledge gap is particularly relevant for Toronto, where rapid expansion of cycling infrastructure has coincided with dramatic growth in the bike share network from 2,854 to over 6,850 stations between 2016 and 2020 [@toronto_bike_share_expansion].
-Understanding this relationship is critical as cities increasingly rely on bike share systems to complement public transit and provide sustainable transportation options.
-
-We estimate the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, examining how protected lanes, on-road lanes, and shared roadways influence system usage.
-Using a difference-in-differences design, we analyze both new installations and infrastructure upgrades while controlling for broader trends in cycling adoption.
-
-Our analysis reveals substantial increases in bike share ridership following infrastructure improvements.
-Post-implementation effects reach 133.1 additional monthly rides per bikeway (p < 0.01), translating to approximately 300,000 additional annual rides across all 186 treated segments - particularly notable as these improvements represent only 16% of the 1,191 bikeways in our study.
-Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides (p < 0.001) - more than double the effect of on-road lanes and triple that of shared roadways.
-These gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment.
+Cycling infrastructure plays a key role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
+While previous research demonstrates that protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
+This knowledge gap is particularly relevant for Toronto, where the city previously committed to significant cycling infrastructure expansion while simultaneously growing its bike share network from 2,854 stations in 2016 to over 6,850 stations in 2020 [@toronto_bike_share_expansion].
+Understanding this relationship is essential as cities increasingly rely on bike share systems to complement public transit and provide sustainable transportation options.
+
+Our study estimates the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, focusing specifically on how different types of infrastructure improvements (protected lanes, on-road lanes, and shared roadways) influence system usage.
+We examine both new installations and upgrades to existing cycling infrastructure to identify treatment effects through a difference-in-differences design.
+By combining detailed ridership data with infrastructure locations, we can isolate the impact of specific improvements while controlling for broader trends in cycling adoption.
+
+Our analysis indicates that infrastructure improvements significantly increase bike share ridership, with effects varying substantially by implementation approach.
+While we observe modest initial increases of 32.4 rides per month during the treatment year, post-implementation effects grow substantially to 133.1 additional monthly rides per bikeway (p < 0.01).
+In aggregate, this translates to approximately 300,000 additional annual rides across all 186 treated segments - a significant impact given that these improvements represent only 12.6% of Toronto's total bikeways.
+Protected bike lanes show particularly strong effects, generating 381.9 additional monthly rides post-implementation (p < 0.001) - more than twice the effect of on-road lanes (169.2 rides, p < 0.001) and triple that of shared roadways (111.7 rides, p < 0.001).
+Notably, these gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment.
 This suggests infrastructure improvements can substantially increase cycling activity even on popular routes.
 
-These findings carry important implications for urban transportation policy.
-Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027 [@cyclingNetworkPlan] presents an opportunity to enhance cycling adoption through strategic implementation.
-Our results challenge conventional assumptions about prioritizing underserved areas, as improvements to high-traffic routes generated significant additional ridership, highlighting the complex trade-offs in infrastructure resource allocation.
+These findings have important implications for urban transportation policy as cities work to expand sustainable mobility options.
+Toronto's commitment to add 100 kilometers of new cycling infrastructure between 2025 and 2027 [@cyclingNetworkPlan] represents a significant opportunity to enhance cycling adoption, particularly if implemented with attention to the differential impacts of various infrastructure types.
+The substantial ridership gains observed from improving just 12.6% of the network suggest that targeted infrastructure investments can have outsized impacts on system utilization.
+Furthermore, our finding that improvements to already-popular routes generated significant additional ridership challenges the assumption that infrastructure investments should prioritize underserved areas, highlighting the complex trade-offs cities face when allocating limited resources for cycling infrastructure.
 
@@
 This study draws on three datasets from the City of Toronto's Open Data Portal [@torontoOpenData]: the Cycling Network (Bikeways) data [@bikeWayData], and both Bike Share Ridership and Stations data [@bikeShareData].
-Together, these datasets provide a comprehensive view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023.
-All data cleaning and analysis was conducted using R [@RCoreTeam2023], with key packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], and `sf` [@sf] for spatial analysis.
+Together, these datasets provide a detailed view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023.
+While alternative cycling metrics exist, such as manual traffic counts and automated counter data, these sources suffer from inconsistent maintenance and ceased updates after 2019.
+Our selected datasets offer better temporal coverage and data quality, with regular updates ensuring current, reliable measurements of both infrastructure changes and system usage.
+
+All data cleaning and analysis was conducted using R [@RCoreTeam2023], with key packages including `tidyverse` [@tidyverse], `dplyr` [@dplyr], `readr` [@readr], `jsonlite` [@jsonlite], `arrow` [@arrow], and `sf` [@sf] for spatial analysis.
 
@@
 
-A crucial methodological choice was the use of relative timing rather than calendar years.
+A vital methodological choice was the use of relative timing rather than calendar years.
 Due to bikeways having different upgrade or installation years, we align their treatment event year using a relative period.
@@
 Our primary outcome variable is seasonally-adjusted monthly ridership for each bikeway (`monthly_rides_adj`).
-@fig-avg-monthly-ridership reveals several important patterns in ridership behavior.
+@fig-avg-monthly-ridership shows several important patterns in ridership behavior.
 The raw ridership (black line) displays pronounced seasonal fluctuations, with consistent peaks during summer months and troughs in winter, reflecting Toronto's strong cycling seasonality.
@@
 
-@fig-infra-type reveals distinct infrastructure patterns between groups.
+@fig-infra-type presents distinct infrastructure patterns between groups.
 While protected lanes dominate both treatment and control groups, comprising approximately 50% of bikeways in each, the distribution of other types differs notably.
@@
 Further categorization into bikeway types, and sub-treatment groups allow us to examine whether certain types of infrastructures and types of changes yield different impacts on ridership patterns.
+
+@tbl-summary-statistics in the Appendix provides detailed quantitative breakdowns of both our outcome variables (showing pre/post/treatment period ridership counts and adjusted means) and treatment variables (showing the distribution of protected lanes, on-road lanes, and shared roadways across treatment and control groups).
 
@@
 
-Our analysis is ran in R [@RCoreTeam2023] using the `fixest` package [@fixest], which provides estimation methods for fixed effects models.
+While Regression Discontinuity Design (RDD) was considered as an alternative approach, our data's temporal granularity (only observing treatment years rather than specific dates) and the gradual nature of infrastructure effects made DiD more appropriate for capturing the full impact of these changes.
+Moreover, the lack of a clear "forcing variable" that determines treatment status and the potential anticipation effects during construction periods would violate key RDD assumptions, making it unsuitable for our analysis.
+
+Our DiD analysis is ran in R [@RCoreTeam2023] using the `fixest` package [@fixest], which provides estimation methods for fixed effects models.
 This package is particularly suitable for our analysis as it handles multiple fixed effects efficiently and provides robust standard errors clustered at the bikeway level.
@@
 ### Parallel Trends Test
-A crucial assumption of the DiD design is that treated and control bikeways would have followed parallel trends in the absence of treatment [@cuborican2024econometrics].
+An important assumption of the DiD design is that treated and control bikeways would have followed parallel trends in the absence of treatment [@cuborican2024econometrics].
 While this counterfactual cannot be directly tested, we can examine pre-treatment trends to assess the plausibility of this assumption.
@@
 We extend our analysis to examine treatment effect heterogeneity across different types of infrastructure changes.
-This investigation is crucial for understanding which interventions are most effective at increasing ridership.
+This investigation is necessary for understanding which interventions are most effective at increasing ridership.
 We estimate two separate models:
@@
 
-These heterogeneity analyses reveal important nuances in the effectiveness of different infrastructure strategies.
+These heterogeneity analyses shows important distinctions in the effectiveness of different infrastructure strategies.
 The particularly strong effect of protected lanes suggests that safety and separation from vehicle traffic may be key factors in encouraging Bike Share usage.
@@
 # Discussion {#sec-discussion}
-Our difference-in-differences analysis of Toronto's cycling infrastructure changes between 2019-2021 reveals nuanced insights about how different types of cycling infrastructure improvements influence Bike Share utilization.
+Our difference-in-differences analysis of Toronto's cycling infrastructure changes between 2019-2021 demonstrates how different types of cycling infrastructure improvements influence Bike Share utilization.
 The results demonstrate that while infrastructure investments broadly increase cycling adoption, the magnitude and timing of these effects vary substantially by implementation approach and infrastructure type.
@@
 
-The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – reveals the true potential of infrastructure improvements once fully integrated into the transportation network.
+The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – shows the true potential of infrastructure improvements once fully integrated into the transportation network.
 This impact from pre-treatment to post-treatment periods aligns with research from the University of Colorado Denver showing that "building safe facilities for cyclists is one of the biggest factors in road safety for everyone" [@MARSHALL2019100539].
@@
 ### Heterogenous Effect Findings
-Our heterogeneous effects analysis reveals crucial insights about which interventions most effectively promote cycling.
+Our heterogeneous effects analysis show which interventions most effectively promote cycling.
 Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment (p < 0.001) – more than twice the impact of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001).
@@
 
-The comparison between upgrades and new installations provides equally important insights for implementation strategy.
+The comparison between upgrades and new installations provides equally important discoveries for implementation strategy.
 Upgrades to existing infrastructure generate more monthly ridership (152.7 adjusted rides), compared to newly-installing bike lanes (115.5 adjusted rides) post-treatment.
@@
 - **User surveys**: Qualitative research could help explain the mechanisms behind the stronger effects of protected lanes and how infrastructure influences mode choice decisions.
-Such research could particularly focus on understanding the "interested but concerned" demographic identified by @nacto2016 as crucial for expanding cycling adoption.
+Such research could particularly focus on understanding the "interested but concerned" demographic identified by @nacto2016 as vital for expanding cycling adoption.
 - **Network effects**: Future studies could examine how the benefits of infrastructure improvements spill over onto connecting routes and affect system-wide ridership patterns.
@@
 - **Neighborhood-level analysis**: Future research could examine how infrastructure effects vary across Toronto's diverse neighborhoods.
-Such analysis would be particularly valuable as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact.
+Such analysis would be particularly beneficial as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact.
 This geographic heterogeneity analysis could also inform equity considerations in infrastructure deployment, ensuring benefits are distributed across the city's diverse communities.
@@
 The clear relationship between infrastructure quality and ridership, particularly the strong effect of protected lanes, provides evidence-based guidance for cities working to encourage sustainable transportation.
-As Toronto continues expanding its cycling network, these insights can help inform investment priorities and implementation strategies to maximize the impact of infrastructure improvements.
+As Toronto continues expanding its cycling network, these findings can help inform investment priorities and implementation strategies to maximize the impact of infrastructure improvements.
 
@@
 Through carefully designed questions about safety perceptions, route preferences, and time-value tradeoffs, we can identify the specific mechanisms through which infrastructure improvements affect ridership patterns.
-Additionally, our sampling of different user types enables analysis of heterogeneous effects across current, former, and potential users in various neighborhoods, providing a comprehensive understanding of how infrastructure changes influence Bike Share adoption.
+Additionally, our sampling of different user types enables analysis of heterogeneous effects across current, former, and potential users in various neighborhoods, providing a extensive understanding of how infrastructure changes influence Bike Share adoption.
 
@@
 
-"Thank you for your time and participation..."Thank you for completing this survey! Your insights will help us better understand how bike infrastructure influences transportation choices in Toronto.
+"Thank you for your time and participation..."Thank you for completing this survey! Your responses will help us better understand how bike infrastructure influences transportation choices in Toronto.
 
@@
 
-As recommended by Stantcheva [-@Stantcheva2023] for testing persistence of effects, we will implement a comprehensive three-month follow-up survey.
+As recommended by Stantcheva [-@Stantcheva2023] for testing persistence of effects, we will implement a three-month follow-up survey.
 The key objective is to assess whether the stated preferences and intentions from the initial survey translate into actual behavioral changes, while also measuring any infrastructure-induced effects on ridership patterns.
@@
 Furthermore, temporal limitations of our three-month follow-up period may inadequately capture seasonal variations in cycling behavior and long-term adaptation to infrastructure changes, while the results may not generalize well to other urban contexts or represent the full diversity of potential users.
-These limitations suggest that while our survey can provide valuable insights into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.
+These limitations suggest that while our survey can provide imporant findings into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.
 

### 2024-12-03 scripts/02-download_data.R
Fixed comment instructions

@@ -25,7 +25,7 @@ library(httr)
 # 1. Go to Open Data Toronto and search up "Bike Share Toronto Ridership Data"
 # 2. Download the bikeshare-ridership-2022 zip file
 # 3. Unzip the file, and unzip "Bike share ridership 2022-11.zip"
-# 4. Import "Bike share ridership 2022-11.csv" into 01-raw_data/02-raw_bikeshare_data
+# 4. Import "Bike share ridership 2022-11.csv" into `data/01-raw_data/01-raw_bikeshare_data`
 ###############################################################################
 
 

### 2024-12-03 paper/paper.qmd
final paper edits

@@
 # Introduction
-Cycling infrastructure plays a key role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
+Cycling infrastructure plays an essential role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
 While previous research demonstrates that protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
@@
 
-Our study estimates the causal effect of cycling infrastructure changes on Bike Share Toronto ridership between 2019-2021, focusing specifically on how different types of infrastructure improvements (protected lanes, on-road lanes, and shared roadways) influence system usage.
-We examine both new installations and upgrades to existing cycling infrastructure to identify treatment effects through a difference-in-differences design.
-By combining detailed ridership data with infrastructure locations, we can isolate the impact of specific improvements while controlling for broader trends in cycling adoption.
+Our estimand is the causal effect of cycling infrastructure improvements on monthly Bike Share Toronto ridership, measured as the difference in average monthly rides between treated and control bikeways before and after infrastructure changes during 2017-2023.
+Specifically, we estimate how the installation of three infrastructure types (protected lanes, on-road lanes, and shared roadways) and two implementation approaches (new installations versus upgrades) affect the number of Bike Share trips originating within 100 meters of improved bikeways.
+Using a difference-in-differences design that combines ridership data with infrastructure locations, we isolate these treatment effects while controlling for broader trends in cycling adoption.
 
@@
 Control bikeways receive randomly assigned pseudo-treatment years (2019, 2020, or 2021) to enable parallel analysis.
-This approach allows us facilitate direct comparison between treatment and control groups while accounting for the temporal distribution of infrastructure changes.
+This approach allows us to facilitate direct comparison between treatment and control groups while accounting for the temporal distribution of infrastructure changes.
 
@@
   ) +
-  facet_wrap(~treatment, scales = "free_y", 
+  facet_wrap(~treatment, 
              labeller = labeller(treatment = c("TRUE" = "Treatment", "FALSE" = "Control")))
@@
 
-Appendix @sec-additional-figures provide a further breakdown on types of bikeways installed and upgraded throughout the years.
+Appendix @sec-additional-figures provides a further breakdown on types of bikeways installed and upgraded throughout the years.
 @fig-bikeway_type_constructed_graph in the appendix shows consistent bikeway installations every year after 2010, with protected lanes most commonly installed.
@@
 @fig-infra-type presents distinct infrastructure patterns between groups.
-While protected lanes dominate both treatment and control groups, comprising approximately 50% of bikeways in each, the distribution of other types differs notably.
+While protected lanes dominate both treatment and control groups, comprising 50% and 54% respectively, the distribution of other types differs notably.
 The treatment group shows a higher proportion of on-road lanes relative to shared roadways, suggesting a deliberate strategy of upgrading less protected infrastructure.

### 2024-12-13 paper/paper.qmd
fixed table formating, removed stars

@@
 #| label: fig-avg-monthly-ridership
-#| fig-cap: "Average ridership by relative months to treatment"
+#| fig-cap: "Average bike share ridership by relative months to bikeway treatment period"
 #| echo: false
@@
     values = c("grey10", "blue"),
-    labels = c("Raw Rides", "Seasonally Adjusted")
+    labels = c("Raw Rides", "Seasonally Adjusted Rides")
   ) +
@@
 #| label: fig-treatment-control-count
-#| fig-cap: "Total Bikeway Count in Control Vs. Treatment Group"
+#| fig-cap: "Total bikeway count in control and treatment group"
 #| echo: false
@@
 #| label: fig-infra-type
-#| fig-cap: "Infrastrucutre Type Distribution between Control and Treatment Group"
+#| fig-cap: "Bikeway type distribution between control and treatment group"
 #| echo: false
@@
 #| label: fig-treatment-type-total
-#| fig-cap: "Number of New-Installations Vs Upgrades in Treatment Group"
+#| fig-cap: "Number of new-installations and upgrades of bikeways in treatment Group"
 #| echo: false
@@
 #| label: fig-parallel
-#| fig-cap: "Parallel Trends Test: Average Monthly Rides in Pre-Treatment Period"
+#| fig-cap: "Parallel Trends Test: Average monthly rides in pre-treatment period"
 #| echo: false
@@
 modelsummary(
-  list("Main DiD Model" = main_did),
-  stars = TRUE,
-  gof_map = c("nobs", "r.squared", "adj.r.squared")
+  list("Main DiD Estimates" = main_did),
+  statistic = "s.e. = {std.error}, p = {p.value}",
+  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  coef_map = c(
+    "period::post:treatment" = "Post-Treatment Effect",
+    "period::treatment:treatment" = "During-Treatment Effect"
+  ),
 )
@@
 
-# Randomly assign treatment
-set.seed(778)
+set.seed(778) # set same seed as model_data implementation script
 
@@
 modelsummary(
-  list("Placebo Test - Random Assignment" = placebo_random_did),
-  stars = TRUE,
-  gof_map = c("nobs", "r.squared", "adj.r.squared")
+  list("Placebo Test Estimates" = placebo_random_did),
+  statistic = "s.e. = {std.error}, p = {p.value}",
+  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  coef_map = c(
+    "period::post:placebo_treatment" = "Post-Placebo Treatment Effect",
+    "period::treatment:placebo_treatment" = "During-Placebo Treatment Effect"
+  ),
 )
@@
 modelsummary(
-  list("Heterogeneous Effects - Bikeway Type" = did_hetero_lane),
-  stars = TRUE,
+  list("Heterogenous Estimates by Bikeway Type" = did_hetero_lane),
+  statistic = "s.e. = {std.error}, p = {p.value}",
   gof_map = c("nobs", "r.squared", "adj.r.squared"),
   coef_map = c(
-  "period::treatment:treatment_infra::Control" = "Treatment x Control",
-  "period::treatment:treatment_infra::Treated-OnRoad" = "Treatment x Treated-OnRoad",
-  "period::treatment:treatment_infra::Treated-Protected" = "Treatment x Treated-Protected",
-  "period::treatment:treatment_infra::Treated-Shared" = "Treatment x Treated-Shared",
-  "period::post:treatment_infra::Control" = "Post x Control",
-  "period::post:treatment_infra::Treated-OnRoad" = "Post x Treated-OnRoad",
-  "period::post:treatment_infra::Treated-Protected" = "Post x Treated-Protected",
-  "period::post:treatment_infra::Treated-Shared" = "Post x Treated-Shared"
-  ),
-  output = "latex"
+  "period::treatment:treatment_infra::Control" = "During-Treatment Effect: Control Bikeways",
+  "period::treatment:treatment_infra::Treated-Shared" = "During-Treatment Effect: Shared Roadways",
+  "period::treatment:treatment_infra::Treated-OnRoad" = "During-Treatment Effect: On-Road Lanes",
+  "period::treatment:treatment_infra::Treated-Protected" = "During-Treatment Effect: Protected Lanes",
+  "period::post:treatment_infra::Control" = "Post-Treatment Effect: Control Bikeways",
+  "period::post:treatment_infra::Treated-Shared" = "Post-Treatment Effect: Shared Roadways",
+  "period::post:treatment_infra::Treated-OnRoad" = "Post-Treatment Effect: On-Road Lanes",
+  "period::post:treatment_infra::Treated-Protected" = "Post-Treatment Effect: Protected Lanes"
+  )
 )
@@
 modelsummary(
-  list("Heterogeneous Effects - Treatment Type" = did_hetero_treat_type),
-  stars = TRUE,
+  list("Heterogeneous Estimates by Implementation" = did_hetero_treat_type),
+  statistic = "s.e. = {std.error}, p = {p.value}",
   gof_map = c("nobs", "r.squared", "adj.r.squared"),
   coef_map = c(
-    "period::treatment:sub_treatment_type::Upgraded" = "Treatment x Upgraded",
-    "period::treatment:sub_treatment_type::Newly-Installed" = "Treatment x Newly Installed",
-    "period::post:sub_treatment_type::Upgraded" = "Post x Upgraded",
-    "period::post:sub_treatment_type::Newly-Installed" = "Post x Newly Installed"
-  ),
-  output = "latex"
+    "period::treatment:sub_treatment_type::Upgraded" = "During-Treatment Effect: Upgraded",
+    "period::treatment:sub_treatment_type::Newly-Installed" = "During-Treatment Effect: Newly Installed",
+    "period::post:sub_treatment_type::Upgraded" = "Post-Treatment Effect: Upgraded",
+    "period::post:sub_treatment_type::Newly-Installed" = "Post-Treatment Effect: Newly Installed"
+  )
 )
@@
 
-# Appendix {.unnumbered}
+# Appendix {-}
 # Idealized Survey Methodology {#sec-ideal-survey}
@@
 #| label: tbl-sample-data
-#| tbl-cap: "Sample of processed dataset for DiD Model"
+#| tbl-cap: "Sample of processed dataset for DiD Model and analysis"
 #| echo: false

### 2024-12-14 paper/paper.qmd
final paper edits

@@
 ---
-title: "Understanding the Causal Impact of Bike Lane Installation on Urban Cycling Adoption"
+title: "Understanding the Causal Impact of Bikeway Installation on Urban Cycling Adoption"
 subtitle: "A Difference-in-Differences Analysis Shows Protected Bike Lanes Triple Bike Share Ridership in Toronto, 2017-2023"
@@
 date-format: long
-abstract: "We examine the causal effect of cycling infrastructure improvements on Bike Share Toronto ridership by analyzing how different types of bikeways influence system usage between 2017-2023.
-Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that infrastructure improvements led to significant increases in ridership, with post-implementation effects reaching 133.1 additional monthly rides per bikeway, translating to approximately 300,000 additional annual rides across all treated segments.
-Protected lanes demonstrate the strongest impact, generating 381.9 additional monthly rides - more than double the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides), while upgrades to existing infrastructure showed more immediate benefits than new installations.
-Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027.
+abstract: "We investigate the causal effects of cycling infrastructure improvements on Bike Share ridership in Toronto between 2017 and 2023.
+Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that treated segments experience an additional 133.1 monthly ride per bikeway, translating to nearly 300,000 additional rides annually, or 40% of all rides on treated bikeways post-improvement.
+Protected lanes generate the largest gains, more than tripling usage compared to shared roadways, while upgrades to existing bikeways outpace the benefits of new installations.Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027.
 These results provide evidence-based guidance for cities working to encourage sustainable transportation through cycling infrastructure development."
@@
 Our analysis indicates that infrastructure improvements significantly increase bike share ridership, with effects varying substantially by implementation approach.
-While we observe modest initial increases of 32.4 rides per month during the treatment year, post-implementation effects grow substantially to 133.1 additional monthly rides per bikeway (p < 0.01).
-In aggregate, this translates to approximately 300,000 additional annual rides across all 186 treated segments - a significant impact given that these improvements represent only 12.6% of Toronto's total bikeways.
-Protected bike lanes show particularly strong effects, generating 381.9 additional monthly rides post-implementation (p < 0.001) - more than twice the effect of on-road lanes (169.2 rides, p < 0.001) and triple that of shared roadways (111.7 rides, p < 0.001).
+While we observe modest initial increases of 32.4 rides per month during the treatment year, post-implementation effects grow substantially to 133.1 additional monthly rides per bikeway.
+In aggregate, this translates to approximately 300,000 additional annual rides across all 186 treated segments, with improvements directly attributable to about 40% of the 1.48 million total rides recorded on these segments in the two years following changes.
+This substantial impact is particularly notable given that these improvements represent only 16% of bikeways in our study.
+Protected bike lanes show particularly strong effects, generating 381.9 additional monthly rides post-implementation - more than twice the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides).
 Notably, these gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment.
@@
 Toronto's commitment to add 100 kilometers of new cycling infrastructure between 2025 and 2027 [@cyclingNetworkPlan] represents a significant opportunity to enhance cycling adoption, particularly if implemented with attention to the differential impacts of various infrastructure types.
-The substantial ridership gains observed from improving just 12.6% of the network suggest that targeted infrastructure investments can have outsized impacts on system utilization.
+The substantial ridership gains observed from improving just 16% of the network suggest that targeted infrastructure investments can have outsized impacts on system utilization.
 Furthermore, our finding that improvements to already-popular routes generated significant additional ridership challenges the assumption that infrastructure investments should prioritize underserved areas, highlighting the complex trade-offs cities face when allocating limited resources for cycling infrastructure.
@@
 This study draws on three datasets from the City of Toronto's Open Data Portal [@torontoOpenData]: the Cycling Network (Bikeways) data [@bikeWayData], and both Bike Share Ridership and Stations data [@bikeShareData].
-Together, these datasets provide a detailed view of Toronto's cycling landscape, capturing infrastructure development from 2001-2023 and ridership patterns from 2017-2023.
-While alternative cycling metrics exist, such as manual traffic counts and automated counter data, these sources suffer from inconsistent maintenance and ceased updates after 2019.
-Our selected datasets offer better temporal coverage and data quality, with regular updates ensuring current, reliable measurements of both infrastructure changes and system usage.
+Together, these datasets provide a detailed view of Toronto's cycling landscape, capturing bikeway infrastructure development from 2001-2023 and ridership patterns from 2017-2023.
+While alternative cycling metrics exist, such as manual and automated cycling traffic data, these sources suffer from inconsistent maintenance and ceased updates after 2019.
+Our selected datasets offer better temporal coverage and data quality, with regular updates ensuring current, reliable measurements of both infrastructure changes and bike share system usage.
 
@@
 When either change occurs, the location, infrastructure type, and year of modification are documented.
-A key limitation of this measurement process is its temporal granularity - while the year of each change is recorded, the specific month and day are not, constraining our ability to pinpoint exact implementation dates.
+A key limitation of this measurement process is its temporal granularity - while the year of each change is recorded, the specific start/end date of construction are not documented, constraining our ability to pinpoint exact implementation dates.
 
@@
 ### Limitations and Potential Biases
-Our analysis faces two primary measurement challenges that could affect causal inference.
-First, the Bike Share system cannot track actual route choices between endpoints.
-To address this, we focus on trips originating from stations within 100 meters of bikeways, assuming proximity indicates usage.
-While this provides a reasonable proxy, it may introduce selection bias by excluding riders who use bikeways despite starting from more distant stations.
-
-Second, our measure of infrastructure changes may suffer from reverse causality bias.
-The city might prioritize upgrades for routes that already show high cycling traffic, potentially overstating the causal effect of infrastructure improvements on usage.
-We attempt to mitigate this through our difference-in-differences design and by examining pre-treatment trends, but this limitation should be considered when interpreting our results.
+Our analysis faces several important limitations and biases that affect both the interpretation of results and their broader applicability.
+First, our temporal measurement of infrastructure changes lacks precision as we observe only the year of implementation, not specific dates.
+This annual granularity likely understates treatment-year effects by combining pre-construction, construction, and post-completion periods within the same temporal window.
+
+A second key limitation stems from the concurrent expansion of Toronto's Bike Share network during our study period.
+The system grew from 2,854 stations in 2016 to over 6,850 stations in 2020, with new stations often strategically placed near improved infrastructure [@toronto_bike_share_expansion].
+While this coordinated deployment reflects real-world policy implementation, it complicates our causal interpretation, as ridership increase could be attributed to improved station accessibility, rather than bikeway infrastructure improvements.
+Our study design helps control for system-wide growth, but cannot fully separate these complementary effects.
+
+Third, our study faces selection bias, as infrastructure improvements may be systematically targeted toward areas with rising cycling demand or changing demographics.
+Toronto's decision to upgrade specific routes likely considers factors like population density, proximity to employment centers, and existing ridership patterns.
+Without controlling for these neighborhood-level characteristics, our treatment effects may partially capture broader urban changes rather than purely causal infrastructure impacts.
+
+Finally, our analysis assumes that bike share rides starting near bikeways indicate infrastructure usage.
+While we determine a 100-meter proximity threshold as a reasonable proxy, we cannot observe actual route choices.
+This limitation may lead us to under-count infrastructure usage by excluding riders who access lanes from more distant starting points, or over-count usage if riders don't ultimately use nearby lanes.
+This measurement challenge is especially acute in dense urban areas where multiple bikeways and stations create complex route options.
+
+These limitations and biases suggest several promising directions for future research, particularly through user surveys that could directly measure route choices and infrastructure preferences.
+Additionally, more granular implementation data and neighborhood-level controls could help isolate infrastructure effects from concurrent changes in the urban environment.
 
@@
 ## Outcome Variables
-Our primary outcome variable is seasonally-adjusted monthly ridership for each bikeway (`monthly_rides_adj`).
-@fig-avg-monthly-ridership shows several important patterns in ridership behavior.
+Our outcome variable is seasonally-adjusted monthly ridership for each bikeway, and @fig-avg-monthly-ridership shows several important patterns in ridership behavior.
 The raw ridership (black line) displays pronounced seasonal fluctuations, with consistent peaks during summer months and troughs in winter, reflecting Toronto's strong cycling seasonality.
+These seasonal patterns remain visible even when using relative months because bikeways in our sample received treatments in different calendar years, but share the same relative month structure (e.g., -24 to +24 months around treatment).
 After applying seasonal adjustments (blue line), these cyclical patterns are smoothed while preserving the underlying growth trends.
@@
 ## Treatment Variables
-Our main treatment variable (`treatment`) indicates whether a bikeway received an infrastructure change (either an upgrade or new installation) during our study period 2019-2021.
+Our main treatment variable indicates whether a bikeway received an infrastructure change (either an upgrade or new installation) during our study period 2019-2021.
 @fig-treatment-control-count presents the distribution of bikeways between treatment and control groups, showing a larger pool of control bikeways to enable robust comparison with treated segments.
@@
 
-To explore treatment effect heterogeneity, we examine two additional dimensions of infrastructure changes:
+To explore treatment effect heterogeneity, we further examine two additional dimensions of infrastructure changes:
 
@@
 This aligns with Toronto's broader policy goal of enhancing cyclist safety through infrastructure improvements [@toronto_council_agenda].
-
+Further categorization into bikeway types allow us to examine whether certain infrastructure types yield different impacts on ridership patterns.
 @fig-treatment-type-total demonstrates a nearly even split in treatment approaches, with 98 new installations and 88 upgrades among the 186 treated bikeways.
-This balanced distribution enables us to examine whether infrastructure expansion (new installations) versus enhancement (upgrades) generates different effects on ridership patterns.
-Further categorization into bikeway types, and sub-treatment groups allow us to examine whether certain types of infrastructures and types of changes yield different impacts on ridership patterns.
-
-@tbl-summary-statistics in the Appendix provides detailed quantitative breakdowns of both our outcome variables (showing pre/post/treatment period ridership counts and adjusted means) and treatment variables (showing the distribution of protected lanes, on-road lanes, and shared roadways across treatment and control groups).
+This balanced distribution enables us to examine whether newly-installed or upgraded bikeways generates different effects on ridership, allowing us to suggest targeted policy implications.
+
+@tbl-summary-statistics in the Appendix provides detailed quantitative breakdowns of both our outcome variables (showing pre/post/treatment period ridership counts and adjusted means) and treatment variables (showing the distribution of protected lanes, on-road lanes, and shared roadways), across treatment and control groups.
 
@@
 While the treatment group consistently shows higher monthly average ridership levels than the control group, the trends between the two groups appear to move in parallel before treatment, particularly in the period from month -20 to month -10.
-We exclude the six months immediately preceding treatment from this visual analysis due to possible overlap in construction, avoiding potential anticipation effects.
 
@@
        aes(x = relative_month, 
-           y = monthly_rides_adj, color = treatment)) +
+           y = monthly_rides_adj, 
+           color = treatment)) +
   stat_summary(geom = "line", fun = mean, size = 1) +
+  geom_smooth(method = "lm", se = FALSE, size = 0.5, alpha = 0.5,
+              linetype = "dashed") +
   labs(
@@
   theme_minimal() +
-  theme(legend.position = "top",
+  theme(
+    legend.position = "top",
     panel.grid.major = element_line(color = "gray85", linewidth = 0.35)
@@
 The model estimates that infrastructure improvements led to a significant increase in Bike Share ridership.
-During the treatment year, we observe a marginally significant increase of 32.4 rides per month (p < 0.10).
+During the treatment year, we observe a marginal increase of 32.4 rides per month (p < 0.1).
 The effect becomes substantially larger in the post-treatment period, with an estimated increase of 133.1 rides per month (p < 0.01), compared to the pre-treatment period.
 
-The high R-squared value of 0.785 indicates that our model explains a substantial portion of the variation in ridership.
+The high R-squared adjusted value of 0.781 indicates that our model explains a substantial portion of the variation in ridership.
 The inclusion of bikeway and time fixed effects helps control for unobserved confounders that might affect ridership patterns.
@@
   statistic = "s.e. = {std.error}, p = {p.value}",
-  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  gof_map = c("r.squared", "adj.r.squared"),
   coef_map = c(
@@
     "period::treatment:treatment" = "During-Treatment Effect"
-  ),
+  )
 )
@@
     "period::treatment:placebo_treatment" = "During-Placebo Treatment Effect"
-  ),
+  )
 )
@@
 ### Heterogeneous Effects
-We examine treatment effect heterogeneity across different types of infrastructure and implementation strategies.
-@tbl-hetero-lane presents the results by bikeway type, with each coefficient representing the change in ridership compared to the pre-treatment period for that specific infrastructure type.
-Protected lanes show the most dramatic impact, generating an additional 381.9 rides per month post-treatment (p < 0.001) – more than twice the effect of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001).
+We examine treatment effect heterogeneity across different types of bikeway and implementation strategies.
+@tbl-hetero-lane presents the results by bikeway type, with each coefficient representing the change in ridership compared to the pre-treatment period for that specific bikeway type.
+Protected lanes show the most impact, generating an additional 381.9 rides per month post-treatment, whereas on-road lanes and shared roadways generated an additional 169.2 and 111.7 additional monthly rides, respectively.
 The substantially larger effect of protected lanes suggests that perceived safety significantly influences riders' willingness to use Bike Share.
@@
 
-@tbl-hetero-treat compares the effectiveness of upgrading existing infrastructure versus installing new bikeways.
-Upgrades to existing infrastructure generate larger and more immediate benefits, showing both a significant treatment-year effect (43.8 additional rides, p < 0.10) and a strong post-treatment effect (152.7 additional rides, p < 0.01).
-In contrast, newly-installed bikeways show no significant effect during the treatment year and a smaller post-treatment increase (115.5 rides, p < 0.10), suggesting that improvements to established cycling routes may be more effective at increasing ridership than creating entirely new routes.
+@tbl-hetero-treat compares the effectiveness of upgrading existing infrastructure against installing new bikeways.
+Our results show that upgrades to existing infrastructure generate larger and more immediate benefits, showing both a significant treatment-year effect of 43.8 additional monthly rides, and a strong post-treatment effect of 152.7 additional monthly rides.
+In contrast, newly-installed bikeways show no significant effect during the treatment year (p = 0.280) and a smaller post-treatment increase (115.5 rides, p < 0.10), suggesting that improvements to established cycling routes may be more effective at increasing ridership than creating entirely new routes.
 
@@
 The evolution of treatment effects over time tells an important story about behavior change and infrastructure adoption.
-During the treatment year, we observe a modest but marginally significant increase of 32.4 rides per month (p < 0.10).
-This relatively small initial effect likely reflects several factors.
+During the treatment year, we observe a modest increase of 32.4 rides per month, likely reflecting several factors.
 First, infrastructure improvements often cause temporary disruption during construction, potentially suppressing immediate usage of bike share in the surrounding areas.
@@
 
-The significant increase in the post-treatment period – to 133.1 additional rides per month (p < 0.01) – shows the true potential of infrastructure improvements once fully integrated into the transportation network.
+The significant increase in the post-treatment period to 133.1 additional rides per month shows the true potential of infrastructure improvements once fully integrated into the transportation network.
 This impact from pre-treatment to post-treatment periods aligns with research from the University of Colorado Denver showing that "building safe facilities for cyclists is one of the biggest factors in road safety for everyone" [@MARSHALL2019100539].
@@
 To put these effects in perspective, our treatment estimate of 133.1 additional monthly rides per bikeway translates to approximately 24,757 additional monthly rides across all 186 treated segments, or nearly 300,000 additional rides annually.
-This substantial increase is reflected in the raw data (see @tbl-summary-statistics in Appendix @sec-additional-figures) - treated segments averaged 331.92 monthly rides post-treatment compared to 209.03 pre-treatment, while control segments saw a more modest increase from 85.59 to 107.77 monthly rides over the same period.
-The magnitude of this effect is particularly notable given that treated segments represent only about 16% of Toronto's total bikeways (186 out of 1,191 total segments) but generated significant ridership gains, suggesting that targeted infrastructure improvements can have outsized impacts on system utilization.
+Given that treated bikeways recorded about 1.48 million total rides in the two years following improvements, our estimates suggest that 40% of these rides can be linked directly to the new or upgraded infrastructure.
+This substantial increase is reflected in the raw data (see @tbl-summary-statistics in Appendix @sec-additional-figures), where treated segments averaged 331.92 monthly rides post-treatment compared to 209.03 pre-treatment, while control segments saw a more modest increase from 85.59 to 107.77 monthly rides over the same period.
+The magnitude of this effect is particularly notable given that treated segments represent only about 16% of total bikeways in our study (186 out of 1,191 total segments) but generated significant ridership gains, suggesting that targeted infrastructure improvements can have outsized impacts on system utilization.
 
@@
 Our heterogeneous effects analysis show which interventions most effectively promote cycling.
-Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment (p < 0.001) – more than twice the impact of on-road lanes (169.2 rides, p < 0.001) and more than triple the effect of shared roadways (111.7 rides, p < 0.001).
+Protected lanes show remarkably strong effects, generating 381.9 additional monthly rides post-treatment, more than twice the impact of on-road lanes, and more than triple the effect of shared roadways.
 This substantial difference in effectiveness likely stems from:
@@
 - **User confidence**: Protected infrastructure may encourage less experienced cyclists to try Bike Share, expanding the user base beyond confident urban cyclists.
-- **Urban density patterns**: Protected lanes are more commonly implemented in denser, high-traffic areas like downtown Toronto, where higher population density and concentration of destinations naturally generate more cycling trips.
+- **Urban density patterns**: Protected lanes are more commonly implemented in denser high-traffic areas like downtown Toronto, where higher population density and concentration of destinations naturally generate more cycling trips.
 This geographic distribution may partially explain their stronger treatment effects compared to other infrastructure types.
@@
 The comparison between upgrades and new installations provides equally important discoveries for implementation strategy.
-Upgrades to existing infrastructure generate more monthly ridership (152.7 adjusted rides), compared to newly-installing bike lanes (115.5 adjusted rides) post-treatment.
-Additionally, during the treatment period, upgrades to existing infrastructure show immediate effect to ridership (43.8 rides, p < 0.10), whereas new installations show no significant treatment-year effect.
+Upgrades to existing infrastructure generate more monthly ridership (152.7 rides), compared to newly-installed bikeways (115.5 rides) post-treatment.
+Additionally, during the treatment period, upgrades to existing infrastructure show immediate effect to ridership, whereas new installations show no significant treatment-year effect.
 This pattern suggests that improving existing cycling routes may be more effective than expanding into new areas, particularly for short-term ridership gains.
@@
 - **Spatial relationships**: While we control for station proximity to bikeways, we cannot observe actual route choices.
-This may lead us to underestimate infrastructure usage if riders access lanes from more distant starting points, or even overestimate usage if riders don't use bike lanes, despite starting near one.
-This limitation is particularly significant in downtown areas where the density of both bike lanes and stations makes it difficult to isolate specific infrastructure effects.
+This may lead us to underestimate infrastructure usage if riders access bikeways from more distant starting points, or even overestimate usage if riders don't use bikeways, despite starting near one.
+This limitation is particularly significant in downtown areas where the density of both bikeways and stations makes it difficult to isolate specific infrastructure effects.
 - **Station placement effects**: Our results may partially reflect strategic placement of new Bike Share stations near improved infrastructure rather than purely infrastructure-driven changes in behavior.
@@
 - **Neighborhood-level analysis**: Future research could examine how infrastructure effects vary across Toronto's diverse neighborhoods.
-Such analysis would be particularly beneficial as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact.
+Such analysis would address our studies selection bias and be particularly beneficial as Toronto expands Bike Share into new areas, helping identify where different types of infrastructure improvements might have the greatest impact.
 This geographic heterogeneity analysis could also inform equity considerations in infrastructure deployment, ensuring benefits are distributed across the city's diverse communities.
@@
 ### Sample Size and Target Audience
-Our survey will be targeted towards: current and former (In-active user for over 6 months) Bike Share members, and non-bike share users living within 200 meters of Bike Share Station.
-Our target sample size will be 3,000 respondents:
+Our survey will be targeted towards: current and former (In-active user for over 6 months) Bike Share members, and non-bike share users living within 200 meters of Toronto Bike Share Stations.
+To ensure our surveys statistical reliability and stays within our budget, it will aim to target 3,000 respondents:
 
@@
 ### Budget Allocation
-Our allocation of a \$100,000 budget, will be:
+With an allocation of a \$100,000 budget, and to ensure we reach our target sample size, we will distribute our budget as follows:
 
@@
 pre_treatment_data <- analysis_data %>%
-  filter(period == "pre" & relative_month < -5) %>%
+  filter(period == "pre" & relative_month < -6) %>%
   mutate(relative_month_treatment_interaction = interaction(
@@
 modelsummary(
-  list("Coefficient (std.error)" = parallel_trends_test),
-  stars = TRUE,
-  gof_map = c("nobs", "r.squared", "adj.r.squared"),
+  list("Estimates" = parallel_trends_test),
   data = pre_treatment_data,
-  estimate = "{estimate}{stars} ({std.error})", 
-  statistic = NULL
+  estimate = "{estimate} (p = {p.value})",
+  statistic = NULL,
+  gof_map = c("r.squared", "adj.r.squared"),
+  coef_map = c(
+    "relative_month::-24:treatment" = "-24 Months to Treatment",
+    "relative_month::-23:treatment" = "-23 Months to Treatment",
+    "relative_month::-22:treatment" = "-22 Months to Treatment",
+    "relative_month::-21:treatment" = "-21 Months to Treatment",
+    "relative_month::-20:treatment" = "-20 Months to Treatment",
+    "relative_month::-19:treatment" = "-19 Months to Treatment",
+    "relative_month::-18:treatment" = "-18 Months to Treatment",
+    "relative_month::-17:treatment" = "-17 Months to Treatment",
+    "relative_month::-16:treatment" = "-16 Months to Treatment",
+    "relative_month::-15:treatment" = "-15 Months to Treatment",
+    "relative_month::-14:treatment" = "-14 Months to Treatment",
+    "relative_month::-13:treatment" = "-13 Months to Treatment",
+    "relative_month::-12:treatment" = "-12 Months to Treatment",
+    "relative_month::-11:treatment" = "-11 Months to Treatment",
+    "relative_month::-10:treatment" = "-10 Months to Treatment",
+    "relative_month::-9:treatment" = "-9 Months to Treatment",
+    "relative_month::-8:treatment" = "-8 Months to Treatment",
+    "relative_month::-7:treatment" = "-7 Months to Treatment"
+  )
 )

### 2024-12-14 paper/paper.qmd
fixed spelling/grammer

@@
 Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that treated segments experience an additional 133.1 monthly ride per bikeway, translating to nearly 300,000 additional rides annually, or 40% of all rides on treated bikeways post-improvement.
-Protected lanes generate the largest gains, more than tripling usage compared to shared roadways, while upgrades to existing bikeways outpace the benefits of new installations.Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027.
+Protected lanes generate the largest gains, more than tripling usage compared to shared roadways, while upgrades to existing bikeways outpace the benefits of new installations.
+Our findings suggest that strategic implementation of protected infrastructure can substantially increase cycling activity, with implications for Toronto's planned addition of 100 kilometers of cycling infrastructure between 2025-2027.
 These results provide evidence-based guidance for cities working to encourage sustainable transportation through cycling infrastructure development."
@@
 ## Survey Objectives
-Following Stantcheva's [-@Stantcheva2023] framework, we recognize that surveys are not merely for data colletion, but rather "creating the process that will generate the data." The goal of our idealized survey is to measure otherwise "invisible factors" regarding cycling infrastructures influence on bike share usage, specifically:
+Following Stantcheva's [-@Stantcheva2023] framework, we recognize that surveys are not merely for data collection, but rather "creating the process that will generate the data." The goal of our idealized survey is to measure otherwise "invisible factors" regarding cycling infrastructures influence on bike share usage, specifically:
 

### 2024-12-15 paper/paper.qmd
fixed spelling/grammer

@@
 abstract: "We investigate the causal effects of cycling infrastructure improvements on Bike Share ridership in Toronto between 2017 and 2023.
-Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that treated segments experience an additional 133.1 monthly ride per bikeway, translating to nearly 300,000 additional rides annually, or 40% of all rides on treated bikeways post-improvement.
+Using a difference-in-differences design with over 7.4 million rides across 1,191 bikeways, we find that treated segments experience 133.1 additional monthly ride per bikeway, translating to nearly 300,000 additional rides annually, or 40% of all rides on treated bikeways post-improvement.
 Protected lanes generate the largest gains, more than tripling usage compared to shared roadways, while upgrades to existing bikeways outpace the benefits of new installations.
@@
 Cycling infrastructure plays an essential role in urban transportation, yet quantifying its impact on cycling adoption remains challenging.
-While previous research demonstrates that protected bike lanes can reduce cycling injuries by up to 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
+While previous research indicates that protected bike lanes may reduce cycling injuries by as much as 89% [@harris2020lane] and increase ridership by up to 171% on specific routes [@nacto2016], less is known about how different types of infrastructure improvements affect bike share system utilization.
 This knowledge gap is particularly relevant for Toronto, where the city previously committed to significant cycling infrastructure expansion while simultaneously growing its bike share network from 2,854 stations in 2016 to over 6,850 stations in 2020 [@toronto_bike_share_expansion].
@@
 Protected bike lanes show particularly strong effects, generating 381.9 additional monthly rides post-implementation - more than twice the effect of on-road lanes (169.2 rides) and triple that of shared roadways (111.7 rides).
-Notably, these gains occurred despite treated bikeways having higher initial ridership at 209.03 monthly rides, compared to control bikeways 85.59 monthly rides pre-treatment.
+Notably, these gains occurred even though treated bikeways started with higher initial ridership (209.03 monthly rides) than control bikeways (85.59 monthly rides) pre-treatment..
 This suggests infrastructure improvements can substantially increase cycling activity even on popular routes.
@@
 
-## Infrastucture Development and Policy Implications
+## Infrastructure Development and Policy Implications
 Our findings have substantial implications for cycling infrastructure development strategies.
@@
 ### Sample Size and Target Audience
-Our survey will be targeted towards: current and former (In-active user for over 6 months) Bike Share members, and non-bike share users living within 200 meters of Toronto Bike Share Stations.
+Our survey will be targeted towards: current and former (inactive user for over 6 months) Bike Share members, and non-bike share users living within 200 meters of Toronto Bike Share Stations.
 To ensure our surveys statistical reliability and stays within our budget, it will aim to target 3,000 respondents:
@@
 Furthermore, temporal limitations of our three-month follow-up period may inadequately capture seasonal variations in cycling behavior and long-term adaptation to infrastructure changes, while the results may not generalize well to other urban contexts or represent the full diversity of potential users.
-These limitations suggest that while our survey can provide imporant findings into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.
+These limitations suggest that while our survey can provide important findings into the relationship between infrastructure and ridership, findings should be interpreted alongside other data sources and methodologies.
 
